### 2024-06-10 docs/verticality_future_mkd.Rmd
starting intro revisions

---
date: "`r Sys.Date()`"
author: "Your Name"
title: "officedown template"
output: 
  officedown::rdocx_document:
    mapstyles:
      Normal: ['First Paragraph']
plots:
  caption:
    style: Image Caption
    pre: 'Figure '
    sep: ': '
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE)
library(officedown)
library(officer)
library(tidyverse)
library(spatialreg)

fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

\newpage

# Introduuction

Vertical stacking of species from beneath the ground to the top of the canopy is an emblematic signature of biologically rich regions and a major driver of biogeographic patterns from local to global scales [@macarthurPopulationEcologyWarblers1958; @scheffersIncreasingArborealityAltitude2013; @oliveiraVerticalStratificationInfluences2019]. In MacArthur's [-@macarthurPopulationEcologyWarblers1958] seminal work on verticality in warblers, he notes distinct partitioning of vertical niche space among species in a community. Since then, vertical stratification at local scales has been widely observed across diverse taxonomic groups [@bashamVerticalStratificationPatterns2023a]. At regional scales, patterns of community verticality vary across environmental gradients, generally increasing in wetter habitats [@bashamVerticalStratificationPatterns2023a; @scheffersIncreasingArborealityAltitude2013]. For example, verticality of frogs increases at higher elevations where cooler and wetter climates increase habitat suitability in the canopy relative to low elevations where hot and dry conditions compress frog communities toward the ground [@scheffersIncreasingArborealityAltitude2013]. Globally, verticality increases toward the warmer and wetter tropics, where vertical partitioning may facilitate high biodiversity in structurally complex forests [@macarthurBirdSpeciesDiversity1961; @oliveiraVerticalStratificationInfluences2019]. Despite widespread agreement that verticality influences patterns of biodiversity across spatial scales, global patterns of verticality and its environmental drivers, including climatic and vegetative factors, have not been comprehensively described for vertebrates.

## possible paragraph describing hypotheses about climate and verticality
- Vertical stratification -> higher richness because species can partition the canopy
- Therefore expect higher richness to be associated with increasing verticality
- However, climate can limits verticality
- We propose that climate limits verticality
- Where climate is suitable, canopy height will positively impact verticality
- where climate is not suitable, canopy height will have no impact on verticality

As climate change causes the reorganization of species assemblages globally [@scheffersBroadFootprintClimate2016], patterns of verticality may be impacted. Changes in taxonomic composition of assemblages changes, are often associated with changes in functional composition. Consequently, widespread patterns of deborealization, tropicalization, and thermophilization have occurred in response to warming temperatures [@feeleyClimatedrivenChangesComposition2020; @lajeunesseTemporalAnalysisGBIF2023; @chustCrossbasinCrosstaxaPatterns2024; @borderieuxExtinctionDrivesRecent2024]. Space-for-time substitutions can be used to predict impacts of climate change on the functional composition of species assemblages. For example, the thermal tolerance of assemblages displays predictable variation across elevation gradients. As the climate warms and species' ranges shift toward higher elevations, the thermal tolerance of assemblages increases as cool-adapted species become locally extirpated and warm-adapted species immigrate into communities (a process known as thermophilization) [@feeleyClimatedrivenChangesComposition2020].

```{r}
birds = read.csv("./../data/derivative_data/species_data/birds_breedresident_spdat_eltonvert_forestsOnly.csv")
mammals = read.csv("./../data/derivative_data/species_data/mammals_spdat_moura_forestsOnly.csv")
amph = read.csv("./../data/derivative_data/species_data/amph_spdat_moura_forestsOnly.csv")
repts = read.csv("./../data/derivative_data/species_data/rept_spdat_moura_forestsOnly.csv")

nbirds = nrow(birds)
nmammals = nrow(mammals)
namph = nrow(amph)
nrept = nrow(repts)

```

Relationships between climate and verticality across spatial and temporal environmental gradients have led to predictions that verticality may decline as climates become warmer and drier (cite). While forest understories are protected from extreme climatic conditions, the canopy experiences larger fluctuations in temperature and humidity [@scheffersTropicalMountainPasses2018; @leahyVerticalNicheElevation2021; @defrenneGlobalBufferingTemperatures2019]. Under climatically stressful conditions, arboreal species may move to lower vertical niche positions, reducing community verticality. Temporary vertical shifts have been observed for decades across a variety of biomes [@macarthurPopulationEcologyWarblers1958; @bashamVerticalStratificationCollapses2020; @kosheleffTemperatureInfluenceActivity2009; @eppleyFactorsInfluencingTerrestriality2022]. For example, arboreal frogs in a Panamanian rainforest move toward the understory during the dry season to escape the hot and dry conditions in the canopy [@bashamVerticalStratificationCollapses2020]. Similarly, arboreal primates from the Americas to Madagascar increase their use of terrestrial habitats under hotter temperatures [@eppleyFactorsInfluencingTerrestriality2022]. Yet how vertical niche shifts may alter global patterns of verticality as the climate warms is not well understood. Here, we integrate range maps of `r nbirds` birds, `r nmammals` mammals, `r nrept` reptiles, and `r namph` amphibians with global climate and vegetation data to examine global patterns and environmental drivers of vertebrate verticality in wooded biomes across the globe. We then examine how climate change may impact environmental conditions that have led to present day patterns of verticality.

# Methods

*infographic for methods*

```{r}
mammals = fread("data/derivative_data/gridcell_data/mammals_comdat/mammals_comdat_parallel_forestsOnly.csv") %>% 
  dplyr::select(vert.mean, vert.mean.ses, p.arb, p.ter, p.fos, biome, ecoregion, realm)

amphibians = fread("data/derivative_data/gridcell_data/amphibians_comdat/amph_comdat_parallel_forestsOnly.csv") %>% 
  dplyr::select(vert.mean, vert.mean.ses, p.arb, p.ter, p.fos, biome, ecoregion, realm)

mammals.arb.tropicalmoist = round(mean(mammals$p.arb[mammals$biome == "Tropical & Subtropical Moist Broadleaf Forests"]) * 100, 2)
```

## Assemblage verticality

We mapped the verticality of vertebrate assemblages in wooded habitats using 50 km grid cells and a cylindrical equal area projection. We identified wooded habitats by ecoregions including forests, woodlands, taiga, várzea, Yungas, savanna, thicket, and mallee as defined by Dinnerstein et al. [-@dinersteinEcoregionBasedApproachProtecting2017]. Range maps were obtained from the IUCN red list for mammals and amphibians [@iucnIUCNRedList2023], BirdLife International for birds [@birdlifeinternationalandhandbookofthebirdsoftheworldBirdSpeciesDistribution2022], and the Global Assessment of Reptile Distributions (GARD) for reptiles [@rollGlobalDistributionTetrapods2017; @caetanoAutomatedAssessmentReveals2022]. For birds, we only used ranges that covered breeding and resident ranges, because seasonality impacts the verticality of assemblages. To calculate the verticality of each assemblage (i.e., grid cell), we assigned each species a verticality score between zero and one and took the average verticality of all species whose ranges overlapped with the grid cell. For mammals, reptiles, and amphibians, verticality scores were based on their use of fossorial (0), terrestrial (0.5), and arboreal (1) habitats as defined in the TetrapodTraits database [@mouraPhylogenyinformedCharacterisationGlobal2024]. This dataset uses phylogenetic imputation to identify traits if trait data is not available in the literature. For species where imputation was necessary, we considered a species to use a given vertical stratum if the imputation probability exceeded 0.7. For birds, we calculated verticality scores based on their use of different foraging strata as identified in the EltonTraits database [@wilmanEltonTraitsSpecieslevelForaging2014]. We classified foraging below the water surface, around the water surface, and on the ground as 0.5, in the understory as 0.667, mid-high canopy as 0.8337, and in or above the upper canopy as 1. If species used multiple vertical habitats, vertical scores were averaged. For example, if a species occurred in terrestrial (0.5) and arboreal (1) habitats, it would receive a verticality score of 0.75. Species that occupied only aerial vertical niches or that did not occur in terrestrial habitats were excluded from the analysis.

Verticality exhibited a strong positive correlation with species richness. To examine spatial patterns of verticality independently of species richness, we claculated the standard effect size (SES) of mean verticality (\$ SES = (x_i - \bar{x})/sd(x)\$, where x is a vector of expected mean verticality scores given random selection of species from a species pool) . We defined species pools as species within all wooded habitats in each biogeographic realm [@dinersteinEcoregionBasedApproachProtecting2017] and generated a null distribution of expected verticality given random chance for each grid cell using 100 simulations.

 To minimize impacts of the positive relationship between verticality and species richness on global patterns, we calculated the standard effect size (SES) of mean assemblage verticality, which determines whether verticality is higher (positive) or lower (negative) than expected by random chance given the species richness in each cell and the vertical strategies of species in each geographic realm.

## Environmental data

We obtained present and future global climate data from CHELSA v2.1 [@kargerDataClimatologiesHigh2021; @kargerClimatologiesHighResolution2017]. Present climate data represents the time period 1981-2010 and future climate data represents an ensemble of climate scenarios for SSP585 for the time period 2071-2100. Climate velocity represents the average rate at which thermal isotherms have shifted between the last glacial maximum and present climate conditions [@sandelInfluenceLateQuaternary2011]. Canopy height was optained from [@langHighresolutionCanopyHeight2023] and vegetaion dnesiy Vegetative complexity was calculated as the product of vegetation density [@crowtherMappingTreeDensity2015] and canopy height [@langHighresolutionCanopyHeight2023]. Elevation data was obtained from *data source*. All environemtnal data was resmapled to the resolution and projection of the assemblage data.

## Analysis

We used spatial generalized linear mixed-effects models (GLMMs) to evaluate the impact of vegetative and climatic paramaters on mean and SES verticality. Spatial models were fit using R package sdmTMB [@andersonSdmTMBPackageFast2024], which fits models using maximum marginal likelihood estimation through template model builder (TMB) and approximates spatial Gaussian random fields using stochastic partial differential equation (SPDE)-based approach [@lindgrenExplicitLinkGaussian2011] as implemented in R-INLA [@lindgrenBayesianSpatialModelling2015]. To implement the SPDE approach, we constructed a mesh using fmesher with a 300 km minimum allowable distance between vertices. We fit five alternative models for each taxonomic group, using the same fixed environmental predictors and varying the form of the random-effects structure. The general model is described as

$$
y_s ∼ 𝜇_{s}, \\
𝜇_s = 𝑓^{-1}(𝜲^{main}_s𝛽 + ⍺_g + ⍵_s), \\
⍵_s ∼ MVNormal(0,𝚺_⍵)
$$

where $y_s$ is the mean or SES verticality at location $s$ defined by x and y coordinates, $𝜇_s$ is the expected mean or SES verticality at all locations, and $𝑓^{-1}$ represpents an inverse link function. $𝜲^{main}_s$ is a vector of main effect covariates, 𝛽 represents a vector of estimated corresponding main effect coefficients. We used a Gaussian distribution for SES verticality, while we used a Beta distribution with a logit link function for mean verticality because values are constratined between 0 and 1. Environmental covariates included canopy height, vegetation density, minimum temperature of the coldest month, maximum temperature of the warmest month, precipitation of the driest month, precipitation of the wettest month, precipitation seasonality, and climate change velocity since the LGM. Fitting models, we log10 transformed climate velocity and precipitation of the driest quarter to increase the linearity of the relationship between predictor and response and scaled all predictor variable to have a mean of zero and a standard deviation of 1. $⍺_g$ represents an optional IID random intercept defined by group $g$ ($⍺_g ∼ Normal(0,𝜎^2_⍺))$. The five alternative models each used a different variable to define the random intercept: biome, realm, biorealm, ecoregion, and no random intercept. $⍵_s$ represents a spatial random field at location $s$, which represents additional spatial variation in the data not explained by the environmental covariates or random intercept. The estimated covariance matrix ($𝚺_⍵$) was constrained by the Matérn covariance function, which defines the decay rate of correlation with distance.

We compared alternative models using five-fold cross validation. Grid cells were randomly assigned to one of five folds. Each model was the run five times, training the model using four of the folds (80% of the data) and testing the model on the withheld fold (20% of the data). After fitting the model using the training data, performance of each training set was calculated as the sum of log-likelihood values for the test data. Alternative models were then compared using the total log-likelihood summed across the five folds for each model structure. Across taxonomic groups, the best performing model contained a random intercept defined by ecoregion along with a spatial random field.

## Add spatially varying coefficient for canopy height
We then compared the best performing model to a model that additionally included a spatially varying coefficient for canopy height because we expect that canopy height may have a weaker impact on verticality in regions where verticality is homogeneously low than in regions were verticality is high.

Using the best performing random-effects structure, we compared three alternative models that differed in their fixed effects structure: the the full model with all fixed effects, the full model without the polynomial term for maximum temperature of the warmest month, and the full model without climate velocity. The base model included vegetative and current climatic predictors (canopy height, vegetation density, minimum temperature of the coldest month, maximum temperature of the warmest month, precipitation of the driest month, and precipitation of the wettest month. All these variables are assumed to directly impact the prevalence of different vertical strategies. The second alternative model included a polynomial term for maximum temperature of the warmest month because verticality increases in the tropics, where it is warmer, but extremely high maximum temperatures may limit arboreal strategies. The third alternative model a

To validate the best performing model, we refit it using all the data and calculated quantile residuals using the DHARMa R package.

\*we examined VIFs for predictor variables, sequentially removed variables until all VIFs were equal to or less than 4. We chose variables based on a combination of which ones had the highest VIF factors and which ones we thought had the most direct impact on physiological limits to arboreality. While, this is slightly higher than the standard VIF cutoff, we thought that it was important to keep both maximum temperature of the warmest month and minimum temperature of the coldest month, as the relative influence of these variables is likely to change across latitudes, with maximum temperatures having a larger impact on verticality in the tropics and minimum temperatures having a larger impact on verticality in temperate and boreal regions. We excluded precipitation seasonality because it was strongly correlated with dry season precipitation, has a less direct impact on physiological limits to verticality than dry season precipitation, and varies between current and future climate conditions than dry season precipitation.\*

To examine potential impacts of climate change on assemblage verticality, we predicted models to future climatic data while holding vegetation variables constant at present-day values due to lack of global predictions of future canopy height and vegetation density. From the future predictions, we subtracted predictions to the data used to train the model and mapped the resulting differences.

We define a mesh, specifying a cutoff of 300 km as the minimum allowed distance between mesh verticies to maintain computation efficiencty while allowing suffcient complexity in the spatial field.

Vegetative and climatic parameters included canopy height, vegetation density, maximum temperature of the warmest month, minimum temperature of the coldest month, precipitation of the driest quarter, precipitation of the wettest quarter, precipitation seasonality, and climate change velocity. All predictors had variance inflation factors below 4. We first identified the optimal random effects structure for the model using five-fold cross validation and the sum of the log-likelihood values across folds. Five random effects structures were compared, with each containing a spatial random field calculated from a 300 km mesh. Four of the five structures contained one additional random intercept: biome, realm, biorealm, or ecoregion. Models fit with a spatial random field and ecoregion as a random intercept consistently performed the best.

Using the optimal random effects structure, we then compared fixed effects using five-fold cross validation and the sum of the log-likelihood values for each fold. Our base model consisted of vegetation parameters (canopy height and vegetation density) and present-day climatic conditions (maximum temperature of the warmest month, minimum temperature of the coldest month, precipitation of the driest quarter, precipitation of the wettest quarter, and precipitation seasonality). We compared this model to one that additionally included climate change velocity, which has previously been shown to correlate with assemblage verticality (cite).

Using the optimal fixed effects structure, we refit the model with all points and validated the model by examining the normality and spatial distribution of quantile residuals. We then projected the models for mean and SES verticality for each taxonomic group to future climatic conditions and mapped the difference between predictions of future and present verticality.

Full models were checked using sanity() function.

We used simultaneous autoregressive (SAR) models with a spatial error term to model the impact of environmental factors on the SES of mean verticality. Environmental variables included mean elevation, vegetative complexity, climate velocity, mean annual temperature, maximum temperature of the warmest quarter, minimum temperature of the coldest quarter, diurnal temperature range, temperature seasonality, annual precipitation, precipitation of the wettest quarter, precipitation of the driest quarter, and precipitation seasonality. To reduce multicollinearity among predictors, we initially applied an automatic selection procedure that eliminated the variable with the highest variance inflation factor (VIF) until all VIFs were below three. This resulted in the loss of both maximum temperature of the warmest month and minimum temperature of the coldest month, while temperature seasonality and diurnal temperature range were retained. Since maximum and minimum temperatures exhibit greater change from present to future climate scenarios and are more likely to directly impact verticality, we substituted these two variables for temperature seasonality and diurnal temperature range. Although this adjustment increased VIF, all VIF remained below five. All selected variables for use in the models included mean elevation, vegetative complexity, climate velocity, maximum temperature of the warmest quarter, minimum temperature of the coldest quarter, precipitation of the wettest quarter, precipitation of the driest quarter, and precipitation seasonality.

Prior to modeling, independent variables were checked for linear relationships with the response variable and log10 transformed if the relationship exhibited a strong right skew. Due to zero values in precipitation of the driest quarter, values were transformed using log10(precipitation + 1). If relationships appeared quadratic, a quadratic term was added to the model. Independent variables were then scaled to a mean of zero and standard deviation of one. To determine the optimal neighborhood weights matrix, we visually examined the distance to which spatial autocorrelation occurred in teh response based on correlograms. We then identified different neighborhood scenarios based on distance from the focal cell. The smallest distance included the eight neighbors immediates surrounding the focal cell and subsequently increased to include successive rings of cells up to the maximum distance at which spatial autocorrelation was present in the response variable. For each distance scenario, we produced three row-standardized neighborhood weights matrices based on distance, inverse distance, and inverse squared distance [@verhoefSpatialAutoregressiveModels2018].

We produced simultaneous autoregressive (SAR) models with a spatial error term using the additive effects of all scaled environmental variables for each neighborhood weight matrix and calculated Moran’s I for each model to determine if spatial autocorrelation remained in the residuals. Models in which spatial autocorrelation remained were discarded. Based on the remaining models, we selected the neighborhood weights matrix as the one used in the model with the lowest AICc. Initial model exploration indicated that spatial autocorrelation was present using larger spatial neighborhoods. To reduce computational demands, we therefore based subsequent model selection on spatial neighborhoods from a radius of one grid cell to seven grid cells around the focal cell. Using the optimal neighborhood weights matrix defined by the lowest AICc, we compared all possible combinations of independent variables using the ‘dredge’ function from the MuMIN R package. Optimal weight matrices used inverse distance squared weighting. We then used a model averaging procedure to estimate coefficients and standard errors for all models within 2 \u0394AICc of the best model [@symondsBriefGuideModel2011ods]. For each model within \u0394AICc of the best model, we calculated the AICc weight (wAICc), which represents the probability that the given model is the best (*cite Burnham and Anderson*). Because wAICc values were similar across candidate models, we used full-model averaging, where coefficients for parameters that do not appear in a model are set to zero. A weighted average of coefficients across all models is then calculated to estimate the impact of the parameter on the response variable. We additionally calculated relative variable importance by summing wAICc in the set of models where the variable occurs (*cite Burnham and Anderson*). To estimate model fit, we calculated pseudo-R^2^ as the squared Pearson correlation between predicted and observed values [@kisslingSpatialAutocorrelationSelection2008].

To estimate changes in mean and SES verticality under future climate scenarios, we predicted the averaged model to future environmental conditions, while climate velocity, elevation, and vegetative complexity remained at the values used in the models of present verticality. Predictions were made using a linear model equation with the coefficients from the averaged SAR model (i.e. the predicted trends without the spatial signal) because it is not possible to estimate the the spatial signal without knowing the values of the response variable. To estimate changes in verticality, we therefore used the same method to predict verticality under present environmental conditions and subtracted these values from predicted future verticality. Negative values indicate declines in verticality and positive values indicate increases in verticality.

# Results

## Global patterns of verticality

#Figures to include
# verticality map
# verticality by ecoregion - could be ciruclar bar plot for each realm or biome and taxonomic group
# coefficient plot
```{r}
load("./../results/sdmTMB_models/amphibians_meanvert.RData")
m.amph.meanvert = m.final
load("./../results/sdmTMB_models/amphibians_sesvert.RData")
m.amph.sesvert = m.final

load("./../results/sdmTMB_models/reptiles_meanvert.RData")
m.rept.meanvert = m.final
load("./../results/sdmTMB_models/reptiles_sesvert.RData")
m.rept.sesvert = m.final

load("./../results/sdmTMB_models/mammals_meanvert.RData")
m.mammals.meanvert = m.final
load("./../results/sdmTMB_models/mammals_sesvert.RData")
m.mammals.sesvert = m.final

load("./../results/sdmTMB_models/birds_meanvert.RData")
m.birds.meanvert = m.final
load("./../results/sdmTMB_models/birds_sesvert.RData")
m.birds.sesvert = m.final

```

# change in verticality map

# Supplementary figures/tables
# estimated coefficients and confidence intervals from the best model for each taxa - include all model parameters: fixed effects, spatial paramters (matern range, marginal spatial standard, deviation), distribution parameters

Verticality exhibited a latitudinal gradient across taxa characterized by high verticality in the tropics and lower verticality toward higher latitudes (\@ref(fig:vert-map-global)). This pattern was largely driven by declines in the proportion of arboreal species toward the poles. Patterns of SES verticality mirrored those of mean verticality and highlight declines in arboreal strategies outside tropical broadleaf forests (\@ref(fig:vert-biome)).

Despite widespread coherence of global verticality patterns among taxa and within biomes, regional variation occurred. Within the tropics, verticality was highest in moist broadleaf forests and lowest in dry broadleaf forests (\@ref(fig:vert-biome)). Mammals exhibited the highest mean verticality in moist forests, particularly in the Amazon, the Congo Basin, Madagascar, and tropical and subtropical Asian forests where an average of `r mammals.arb.tropicalmoist`% of species per assemblage exhibited arboreal strategies while the proportion of terrestrial species declined. Verticality of birds closely mirrored this pattern except in Madagascar where SES of mean verticality declined. While reptiles and amphibians displayed similar verticality patterns, there were notable differences. Verticality of reptile assemblages peaked in the rainforests of Indonesia and Madagascar and declined in the neotropics where the proportion of fossorial species increased *map proportion arboreality and fossoriality*. Amphibian verticality additionally increased in the Brazilian Atlantic Forest while declining in SE Asia relative to other vertebrate assemblages.

In temperate biomes, verticality was similar across forests and declined in less dense wooded habitats (e.g., savannas) (\@ref(fig:vert-biome)). Variation in the proportion of species exhibiting fossorial strategies drove differences in patterns of assemblage verticality among vertebrate classes. Amphibians exhibited the highest proportion of fossorial strategies in temperate biomes, peaking in savannas and savanna-woodland transition zones (\@ref(fig:vert-map-global)L, \@ref(fig:vert-biome)). However, temperate rainforests (e.g., the Pacific-Northwest of North America) were notable exceptions, exhibiting higher verticality than than expected by chance. Mammals exhibited a similar pattern, though the proportion of fossorial species increased to a lesser extent (\@ref(fig:vert-map-global)F). In contrast, the proportion of fossorial reptiles declined in temperate regions. Low reptile verticality was therefore driven by declines in the proportion of arboreal species and increases in the proportion of terrestrial species. Though no birds exhibit fossorial strategies, a reduction in the proportion of arboreal species still produced a steep latitudinal gradient in the SES of verticality.

# Environmental drivers of verticality

Models of mean and SES verticality indicated some similarities between the environmental drivers of mean and SES verticality across taxonomic groups, yet noticeable differences remained.
We hypothesized that canopy height would have an overall positive effect on verticality, but that its effect would vary over space, being stronger in tropical regions and lower in temperate regions, which generally have few arboreal vertebrates despite the presence of tall canopies. All of the best models included a spatially varying coefficient for canopy height. Overall, canopy height exhibited a positive effect on verticality of birds and mammals and weak positive effect or no effect on verticality of reptiles and amphibians (\@ref(fig:coef-plt). Over space the impact of canopy height was generally strongest in the Andes and Madagascar

If there is a positive relationship between canopy height and verticality, then taller canopies have more arboreal species
If there is a negative relationship between canopy height and verticality, then shorter canopies have more arboreal species
If there is not relationship between canopy height and verticality, then height of the canopy doesn't matter, potentially because climatic or evolutionary factors determine verticality.

Across taxa, precipitation of the driest and wettest months positively influenced the SES of verticality, though the impact of dry season precipitation was noticeably stronger. 
Impacts of climate on mean verticality generally mirrored those on the SES of verticality, though effects were smaller relative to the impact of canopy height.

variable effects of climate and vegetation across taxa. Mean verticality of birds and mammals was most strongly influenced by canopy height, though responses to canopy height varied across space. Though climatic variables had

Overall, the mean and SES of verticality exhibited similar responses to environmental variables across taxonomic groups. Verticality was positively associated with precipitation of the driest and wettest months and exhibited a quadratic relationship with maximum temperature of the warmest month for mean and SES verticality of birds and reptiles and SES verticality of mammals. The quadratic relationship was characterized by an initial increase in verticality followed by a decline in verticality at extremely high maximum temperatures. Mean verticality of mammals exhibited a relatively small positive relationship with maximum temperature and amphibians  except for amphibians, for which SES of verticality was negatively associated with maximum temperatures. Minimum cold temperatures positively impacted verticality of mammals, such that verticality increased as minimum temperatures increased, while having little impact on birds and mammals. Impacts of climate velocity also varied, negatively impacting birds, mammals, and amphibians, while having a small positive effect on reptiles.

*For difference section - current geographic niche underfilling and future potential to fill unused geographic space by filling currently unoccupied volume in the canopy as climate permits these habitats to become more suitable*

Though arboreal habitats remain in temperate biomes, the entire habitat volume is not occupied. This could be caused by arboreal habitats being climatically unsuitable in temperate regions and lead to a breakdown of the relationship between richness and volume. Overall, these patterns indicate a decline in the volume of habitable space toward the poles despite the structural presence of habitat.

-

    ## climate velocity represents climate instability across geographic timescales and negatively impacted verticality across vertebrate classes

-   Elevation

-   Vegetative complexity - selective logging could negatively impact verticality by reducing structural complexity

-   Decline in verticality in temperate regions despite the avialability of arboreal habitat space indicates physiological challenges of arboreal strategies in seaosnally cold and/or dry environmentals (Oliveira and Scheffers 2019)

-   for some vertebrate classes, fossoriality offers refuge from climate extremes

*plot richness by area x structural complexity and color by latitude* *then plot richness x latitude* *consider adding correlation of verticality with species richness and discuss here*

```{r}
load("./../results/amphibians_sar_vertmean2.RData")
amph.avgmod = vert.avg
amph.summ = summary(amph.avgmod)
amph.summ = round(amph.summ$coefmat.full, 2)
colnames(amph.summ) = c("Estimate", "SE", "z", "p")

load("./../results/mammals_sar_vertmean2.RData")
mammals.avgmod = vert.avg
mammals.summ = summary(mammals.avgmod)
mammals.summ = round(mammals.summ$coefmat.full, 2)
colnames(mammals.summ) = c("Estimate", "SE", "z", "p")

load("./../results/reptiles_sar_vertmean2.RData")
rept.avgmod = vert.avg
repts.summ = summary(rept.avgmod)
repts.summ = round(repts.summ$coefmat.full, 2)
colnames(repts.summ) = c("Estimate", "SE", "z", "p")

load("./../results/birdselton_sar_vertmean.RData")
birds.avgmod = vert.full
birds.summ = summary(birds.avgmod) # make sure spatialreg is loaded for summary to produce correct results
birds.summ = round(birds.summ$Coef, 2)
colnames(birds.summ) = c("Estimate", "SE", "z", "p")

```

# Environmental drivers of verticality

## Climate

Global patterns of verticality could largely be explained by environmental drivers. To examine the impacts of climate, topography, and vegetation on global patterns of verticality, we used simultaneous autoregressive (SAR) models with a spatial error term. The best models explained explained 88-98% of the global variance in the mean and SES of mean verticality across classes (\@ref(fig:model-coefs)). However, the spatial signal in all models was strong (lambda \> 0.9), which resulted in relatively small impacts of environmental variables as well as a lower pseudo-R^2^ value when calculated based only on the predicted trends without the spatial signal.

Precipitation of the driest quarter and minimum temperatures of the coldest quarter had the largest effects on the mean and SES of verticality across vertebrate classes, both exhibiting positive effects barring the impact of minimum temperatures on amphibians, which was insignificant ((\@ref(fig:model-coefs), \@ref(tab:mean-vert-mods), \@ref(tab:sesvert-mods)). Maximum temperatures of the warmest quarter had opposing effects on endotherms and ectotherms, negatively impacting verticality of reptiles and amphibians while generally having positive or insignificant effects on birds and mammals, except at the highest temperatures where verticality declined ((\@ref(fig:model-coefs), \@ref(tab:mean-vert-mods), \@ref(tab:sesvert-mods)). Arboreality is therefore primarily constrained by cold minimum temperatures and limited precipitation during the dry season, while high maximum temperatures additionally limit arboreal strategies in ectotherms. The vertical climate gradient from beneath the ground to the top of the canopy drives these global constraints on arboreality. Greater climatic fluctuations in the canopy coupled with water limitation pose physiological challenges to arboreal lifestyles. For example... In contrast, terrestrial and fossorial habitats are buffered from climate extremes, which facilitates species persistence where hot, cold, or dry conditions may prevent the occurrence of arboreal lifestyles.

Dry conditions negatively impact arboreal lizard, particularly under simultaneously hot environments. <https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecm.1301>

```         
- Though arboreal habitats can exhibit average temperatures over 2C higher than terrestrial habitats (Scheffers et al. 2013) critical thermal maxima of arboreal species may not substantially exceed that of terrestrial species (Carilo Filho et al. 2021)
```

## Historical Climate - climate velocity

Verticality is also generally higher in regions that have experienced historically stable climatic conditions, as indicated by the significant negative relationships between climate velocity and the mean verticality of birds, mammals, and amphibians and SES of verticality for mammals, as well as the absence of positive relationships for any taxa. Low climate velocity (i.e., climate stability) generally increases speciation rates and decreases extinction rates (Colville et al. 2020). In tropical forests, complex vegetation structure characterized by multilayered forests (Terborh 1985) increases opportunities for diversification among arboreal niches (see Colville et al. 2020 11-13). For example, arboreal lineages of reptiles and frogs generally exhibit higher diversification rates [@liWhatDrivesDiversification2022; @moenMicrohabitatClimaticNiche2017], while fossorial strategies have been associated with reduced speciation and increased extinction rates in snakes [@cyriacDiggingTheirOwn2018].

-   climate stability may therefore provide the conditions necessary for diversification in regions where climatic conditions are suitable for arboreal strategies, while increased dspersal capacity of arboreal species throughout 3D arboreal environments facilitates persistence in climatically unstable regions (Sandel et al. 2011; Scheffers et al. 2017).

```{r}
bird.meanvert = load("./../results/sar_mods_forestOnly_forestSES/birds/birds_sar_meanvert.RData")
bird.meanvert = pred.df
bird.meanvert$vertvar = "mean"
bird.sesvert = load("./../results/sar_mods_forestOnly_forestSES/birds/birds_sar_sesvert.RData")
bird.sesvert = pred.df
bird.sesvert$vertvar = "ses"
bird.pred = bind_rows(bird.meanvert, bird.sesvert) %>% mutate(class = "bird")

mammals.meanvert = load("./../results/sar_mods_forestOnly_forestSES/mammals/mammals_sar_meanvert.RData")
mammals.meanvert = pred.df
mammals.meanvert$vertvar = "mean"
mammals.sesvert = load("./../results/sar_mods_forestOnly_forestSES/mammals/mammals_sar_sesvert.RData")
mammals.sesvert = pred.df
mammals.sesvert$vertvar = "ses"
mammals.pred = bind_rows(mammals.meanvert, mammals.sesvert) %>% mutate(class = "mammals")

rept.meanvert = load("./../results/sar_mods_forestOnly_forestSES/reptiles/reptiles_sar_meanvert.RData")
rept.meanvert = pred.df
rept.meanvert$vertvar = "mean"
rept.sesvert = load("./../results/sar_mods_forestOnly_forestSES/reptiles/reptiles_sar_sesvert.RData")
rept.sesvert = pred.df
rept.sesvert$vertvar = "ses"
rept.pred = bind_rows(rept.meanvert, rept.sesvert) %>% mutate(class = "rept")

amph.meanvert = load("./../results/sar_mods_forestOnly_forestSES/amphibians/amphibians_sar_meanvert.RData")
amph.meanvert = pred.df
amph.meanvert$vertvar = "mean"
amph.sesvert = load("./../results/sar_mods_forestOnly_forestSES/amphibians/amphibians_sar_sesvert.RData")
amph.sesvert = pred.df
amph.sesvert$vertvar = "ses"
amph.pred = bind_rows(amph.meanvert, amph.sesvert) %>% mutate(class = "amph")

pred = bind_rows(bird.pred, mammals.pred, rept.pred, amph.pred)

```

# Impacts of climate change on verticality

To examine the potential impacts of climate change on the verticality of vertebrate assemblages, we used the coefficients from the SAR models to predict assemblage verticality for the time period 2071-2100 using the SSP585 future climate scenario, which represents the upper boundary of predicted warming under a scenario of fossil-fueled development. Though climate change will impact vegetation structure, we were unable to incorporate predictions of future vegetation into the models due to a lack of global future predictions of vegetation height and density. Because we could not predict the spatial signal in verticality, we compared predictions under future climate conditions to predictions under present climate conditions that do not incorporate the spatial signal. - We find that x% of communities could experience declines in verticality. - Declines in verticality are concentrated in the tropics. - Between 23 and -23 latitude, x% birds, x% mammals, x% reptiles, and x% amphibians exhibit declines in verticality - Declines in the Amazon were notably severe

## discussion of climate change

-   What causes changes in verticality - vertical niche shift, immigration, or local extinction?

-   in snakes, it is more common for species to exhibit shift from nonfossorial to fossorial than the other way around (Cyriac and Kodondaramiah 2018) - suggests increase in verticality for reptiles could be caused by extinction of fossorial species rather than vertical niche shift higher

    -   further evidence by Cyriac and Kodandoramaiah 2017 - high extinction rates of fossorial snakes during prolonged environmental fluctuations

-   but these patterns aren't consistent - e.g., in lizards, transitions from fossoriality to nonfossoriality is not less common that transisitons in the other direction

-   In temperate regions, we find...

-   Predicted changes in verticality led to homogenization of vertical strategies globally.

# Discussion

In the tropics, the use of arboreal habitat likely contributes to high species richness (cite Oliveira, also see Naikatini et al. 2022, MacArthur). The large volume of habitat in tropical biomes facilitates species packing within a given area as well as expansion across vertically stratified resources. However, habitat volume and structural diversity is generally not related to species richness in forested ecosystems in the US, except for birds (La Rue et al. 2023). Low arboreality toward higher latitudes may explain the breakdown in the volume-richness relationship and be caused by low climatic suitability of arboreal habitats in temperate climates, except for birds which continue to use arboreal habitats at northern latitudes.

-increasing arboreality at norther latitudes suggests that climate change could lead to arboreal habitats in boreal regions becoming more suitable for species, thereby releasing competitive pressure as species shift poleward by providing more habitat volume. -niche attrition - land use change eliminates the volume of arboreal habitats, which leads to declines in arboreality. Will climate change cause further attrition of arboreal habitats?? - elevator to extinction for mountaintop species = digging a grave

-   vertical niche shifts occur in response to novel habitat, global change, etc.
-   e.g., shift in birds to arboreal foraging on islands (Lapiedra et al. 2021 Niche shifts after island colonization...)

Despite the presence of biomes with vertical structure in temperate forests the proportion of arboreality steeply declines, suggesting that the geographic niche space is underfilled by the animal community.

Low climatic suitability may limit arboreality and thus diversity in temperate biomes.

## The influence of vegetation
Indeed, it is expected that the volume of habitat space explain variation in species richness among forested biomes. However, this relationship is not upheld for birds and mammals (LaRue et al. 2023). Declining arboreality in temperate forests that have the vegetation structure to support arboreal strategies may explain this pattern and be caused by environmental factors limiting the suitability of canopy habitats.

# Methods

# Figures

```{r fig.id="vert-map-global"}

```

```{r fig.id="model-coefs"}

```

```{r fig.id="vert-difs-global"}

```

```{r fig.id="vert-homogenization"}

```

# Extended Data Tables

```{r tab.id = "mean-vert-mods", tab.lp="ed", tab.cap="Simultaneous autoregressive model results for models of mean verticality of vertebrate assemblages as a function of climatic, topographic, and vegetative factors."}

```

```{r tab.id = "sesvert-mods", tab.lp="ed", tab.cap="Simultaneous autoregressive model results for models of the standard effect size of mean verticality of vertebrate assemblages as a function of climatic, topographic, and vegetative factors."}

```

# Extended Data Figures

```{r fig.id="vertses-hist-biome", fig.lp="ed", fig.cap="Histograms of verticality by biome for each vertebrate class. Red dotted line indicates the mean verticality for each biome/class."}
# plot SES vert by biome
("figures/supp_figs/vert_biome_forestOnly_forestSES.R")
plt_vertses_biome

```

```{r fig.id="vert-forest", fig.lp="ed", fig.cap="Mean verticality, standard effect size (SES) of mean verticality, and the average proportion of species exhibiting arboreal, terrestrial, and fossorial strategies across latitudinal bands. In maps of mean veticality, assemblages composed of only fossorial species are zero (red) and those composed of only arboreal species are quantified as one (purple). In maps of the SES of mean verticality, negative numbers (red) indicate assemblages have lower verticality than expected by random chance and positive numbers (purple) indicate assemblages have higher verticality than expected by random chance. Line graphs show the average proportion of species exhibiting arboreal (blue), terrestrial (yellow), and fossorial (red) strategies. Proportions for each assemblage can exceed one because species can exhibit multiple vertical strategies.}

```

```{r fig.id="forest-model-coefs"}

```

## Reference

## Acknowledgements

We would like to thank Roger Bivand for useful insights regarding predictions of SAR models.