### 2025-07-09 manuscript.Rmd
text and fig2 revisions

---
date: "`r Sys.Date()`"
author: "Your Name"
title: "Forest canopies are better generators of biodiversity in the tropics"
output: 
  officedown::rdocx_document:
    mapstyles:
      Normal: ['First Paragraph']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE)
library(officedown)
library(officer)

fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)
```

```{r}
# analysis scripts
# storing all figures in figures/ms_figures/
# other figures are from older versions

"scripts/03_analyses/01_canopyheight_richness_relationships.R" # canopy height richness models and plots
```

\newpage

"The observer who leaves the center of the earth by an infinitely small amount compared to the radius can reach a new world, so to speak, and he can observe more variations in the aspect of the soil and more modifications in the atmosphere than he would if he were to pass from one latitude to another" - Humbolt pg 77 Essay on the Geography of Plants

"Current hypotheses for large-scale geographical patterns of species diversity fail to explain the extraordinarily high biodiversity of mountain regions, particularly in the tropics (2–5). We call this challenge “Humboldt’s enigma.”" - Rahbek

For centuries, ecologists have been intrigued by the role of vertical landscape features in biogeographic patterns from local to global scales [@vonhumboltEssayGeographyPlants2013; @macarthurBirdSpeciesDiversity1961; @bashamVerticalStratificationPatterns2023a; @xingEcologicalPatternsProcesses2023a]. In 1799, Alexander von Humbolt began an expedition that would ultimately canonize the role of mountains in global richness patterns [@vonhumboltEssayGeographyPlants2013; @HumboldtsLegacy2019]. [Yet until recently, hypotheses failed to explain the extraordinary richness of mountains particularly in the tropics.] The vertical nature of mountains provides more land surface per unit area than flat regions and produces steep climate gradients over short distances that allow rapid rates of species turnover. Combined, these factors allow mountains to generate an extraordinarily high biodiversity for a given area relative to flat and climatically homogeneous lowlands [@rahbekHumboldtsEnigmaWhat2019]. Yet mountains are better generators of biodiversity in the tropics because high latitude mountains have lower elevation snow lines that truncate the area of land suitable for diverse communities, and greater seasonal temperature variation, which reduces the climatic distinctness of elevational bands relative to tropical mountains [@rahbekHumboldtsEnigmaWhat2019; @vonhumboltEssayGeographyPlants2013].

NEED TO MENTION HUMBOLT'S ENIGMA SOMEWHERE IN FIRST PARAGRAPH AND THAT THE MAIN EXPLANATIONS ARE TOPOGRAPHY AND CLIMATE

Over a century later, Robert MacArthur highlighted the role of another vertical dimension in promoting high biodiversity: the vertical structure of vegetation, which is increasingly being recognized as critical to our understanding of biogegoraphic patterns [@macarthurBirdSpeciesDiversity1961]. Tall, multi-layered forests generate biodiversity through similar mechanisms to mountains. By increasing habitat volume and producing vertical zones of microclimate and resource availability, tall forests allow species to vertically stratify habitat akin to multi-floored apartments [@bashamVerticalStratificationPatterns2023a; @xingEcologicalPatternsProcesses2023a; @macarthurLimitingSimilarityConvergence1967; @macarthurBirdSpeciesDiversity1961; @larueDiversityVolumeRelationships2023]. Thus, taller forests increase niche opportunities, which may allow additional species to join the assemblage and lead to greater richness within a given planimetric area compared to regions with shorter vertical vegetation structure [@xingEcologicalPatternsProcesses2023a; @macarthurBirdSpeciesDiversity1961; @macarthurLimitingSimilarityConvergence1967]. This vertical stacking of species has been described as a mechanistic link between canopy height and species richness, and is thought to contribute to exceptionally high biodiversity in the tropics where canopies are often tall and structurally complex [@oliveiraVerticalStratificationInfluences2019; @rollLinkingVertebrateSpecies2015; @fengForestCanopyHeight2020]. Yet positive impacts of canopy height on richness are not consistent globally, with often weaker relationships in high latitude regions where forests of comparable vertical profiles to those in the tropics harbor fewer species [@larueDiversityVolumeRelationships2023; @macarthurBirdSpeciesDiversity1961; @rollLinkingVertebrateSpecies2015; @fengForestCanopyHeight2020]. - a vertical dimension to Humbolt's enigma

Despite decades of ecological inquiry into the vertical dimensions of habitat structure and biodiversity [@macarthurBirdSpeciesDiversity1961; @scheffersIncreasingArborealityAltitude2013; @xingEcologicalPatternsProcesses2023a], the mechanisms that help or hinder the capacity for tall forests to generate biodiversity remain poorly understood yet are critically important for understanding biogeographic patterns, impacts of climate and landuse change on biodiversity, and the conservation value of forests. Following upon explanations to Humbolt's enigma [@rahbekHumboldtsEnigmaWhat2019], we fill this gap by examining how vegetative complexity and climate may play key roles in mediating the relationship between canopy height and species richness by impacting the capacity for tall canopies to support vertically stratified and species rich arboreal assemblages. For example, the denser layering of vegetation in tropical forests may provide more niche opportunities, allowing tall canopies to support more biodiversity than canopies of equivalent height in temperate regions [CITE or CONCEPT FIG]. Moreover, densely layered vegetation produces vertical microclimate gradients via interactions with macroclimate conditions, solar radiation, and wind speed that are characterized by humid and thermally stable understories to dry and thermally variable canopies that experiences both hot and cold temperature extremes. Physiological constraints imposed on arboreal life strategies by these more extreme conditions, combined with the distinctness of vertical climate bands may impact species turnover from the ground to the canopy and thus the capacity for taller canopies to support increasing richness [@bashamVerticalStratificationCollapses2020] [CITE]. Yet only recently have we been able to address these hypotheses, as our ability to measure vegetation structure and quantify its use by animals globally [AND MODEL MICROCLIMATES GLOBALLY] increased with the collation of large ecological databases and the launch of satellite missions to measure forest height and structural complexity [@popkinSpaceLaserMap2018; @mouraPhylogenyinformedCharacterisationGlobal2024a].

In this study we combine range maps of nearly 24,000 terrestrial vertebrate species and comprehensive databases of vertical habitat use, global macroclimate and forest structure data, and process-based microclimate models to examine how forest structure and climate may mediate the capacity for tall forests to support species rich assemblages via impacts on the vertical structure of vertebrate communities. We first investigated the capacity for taller forests to support increasing species richness in forests and less densely vegetated wooded biomes across latitudes. If the vertical structure of species assemblages mediates the relationship between canopy height and species richness, then we expect positive correlations between the verticality of assemblages (i.e., their average vertical niche position) and species richness. [AND THAT CANOPY HEIGHT WILL EXPLAIN LITTLE VARIATION IN RICHNESS ONCE VERTICALITY HAS BEEN ACCOUNTED FOR]

and expected that tropical canopies would exhibit the fastest rates of increase in richness with canopy height. We then examined the mechanistic processes driving this pattern.  EXPLANATION OF TETRAPODTRAITS AND ELTONTRAITS

, including TetrapodTraits and EltonTraits, to map species richness and the vertical structure of species assemblages in wooded biomes across the globe.

To fill this gap, we combine range maps of nearly 24,000 terrestrial vertebrate species with the TetrapodTraits and EltonTraits databases to map species richness and the vertical structure of species assemblages in wooded biomes across the globe. We first investigated the capacity for taller forests to support greater species richness in forests and less densely vegetated wooded biomes across latitudes and expected that tropical canopies would exhibit the fastest rates of increase in richness with canopy height. We then examined the mechanistic processes driving this pattern. If vertical stratification contributes to the capacity for tall forests to support diverse assemblages, we expect that species rich assemblages will exhibit high verticality (i.e., a high average vertical niche position) and that environmental factors limiting verticality will limit species richness. To test this hypothesis, we examined correlations between assemblage verticality and species richness. We then modeled assemblage verticality of the four vertebrate classes by macroclimate and vegetative structure and investigated the microclimatic processes that underlie these macroclimatic patterns. While three-dimensional climate data is not currently available globally given the large computational requirements of resolving micrometeorological processes within forest canopies, we modeled vertical microclimate profiles in six sites across the globe within tropical, temperate, and boreal forests by integrating forest structure data derived from NASA's Global Ecosystem Dynamics Investigation (GEDI) mission with high temporal resolution macroclimate data in a process-based microclimate model.

# Tall canopies generate biodiversity, but are better generators of biodiversity in the tropics

```{r}
source("scripts/03_analyses/01_canopyheight_richness_relationships.R")
# ratedifs provides table of differences in richness between forests that are 15 and 25 m tall for each biome/taxa combo
```

To evaluate the capacity for tall canopies to generate biodiversity, we modeled vertebrate richness as a function of canopy height for all vertebrates and each vertebrate class globally. We then mapped the residuals of these models to determine where canopy height over- or under-predicted species richness. We then produced separate models for tropical, temperate, and boreal regions in forests and more sparsely vegetated habitats, including woodlands, savannas, and shrublands, to examine variation in the capacity for wooded habitats of the same height to generate biodiversity and variation in the rate of increasing richness with a given increase in canopy height across different regions.

Globally, taller trees, which increase habitat volume, support more species (Fig. S\@ref(fig:global-chrich)). However, global models of vertebrate richness by canopy height overpredict richness at high latitudes (negative residuals) and underpredict richness in the tropics (positive residuals) (Fig. \@ref(fig:fig1), S\@ref(fig:resids-taxa)). Therefore, canopies of equivalent height generally harbor more species in the tropics than in temperate and boreal regions. Furthermore, increasing canopy height yields greater gains in richness in the tropics than at higher latitudes (Fig. \ref(fig:fig1)). For example, an increase in canopy height from 15 m to 25 m increases richness by `r round(ratedifs$dif[ratedifs$biome2 == "Tropical forest" & ratedifs$taxa=="Birds"])` birds and `r round(ratedifs$dif[ratedifs$biome2 == "Tropical forest" & ratedifs$taxa=="Amphibians"])` amphibians in tropical forests and by only `r round(ratedifs$dif[ratedifs$biome2 == "Boreal" & ratedifs$taxa=="Birds"])` birds and `r round(ratedifs$dif[ratedifs$biome2 == "Boreal" & ratedifs$taxa=="Amphibians"])` amphibians in boreal regions.  Within the tropics, increasing canopy height generally yields greater gains in richness in forests than in drier woodlands, savannas, and shrublands, though baseline richness in areas with short vegetation are similar (Fig. \@ref(fig:fig1)). For the same increase in canopy height, the increase in bird richness in tropical forests exceeds the increase in bird richness in tropical woodlands, savannas, and shrublands by `r round(ratedifs$dif[ratedifs$biome2 == "Tropical forest" & ratedifs$taxa=="Birds"]) - round(ratedifs$dif[ratedifs$biome2 == "Tropical woodland, savanna, and shrubland" & ratedifs$taxa=="Birds"])` species. The difference in the capacity for tall canopies to generate biodiversity between densely and sparsely vegetated tropical regions is particularly evident in Madagascar where the global model underpredicts vertebrate richness in the humid forests of the east coast and overpredicts richness in the drier woodlands and shrublands of the southwest (Fig. \@ref(fig:fig1)).

Minor deviations from these patterns may indicate variations in evolutionary drivers of species richness patterns.
  [- This pattern is particularly stark in the western hemisphere. In the eastern hemisphere, the pattern is weaker where evolutionary forces may drive overprediction of richness in parts of the lowland African tropics and Indo-Malaysian. In Indo-Malaysia, island size may reduce vertebrate richness relative to expectations by canopy height.]

However, the described patterns are broadly consistent among taxa and mirror findings that for a given area, mountains have higher richness in the tropics [@rahbekHumboldtsEnigmaWhat2019]. Thus, increases in habitat volume via either mountains or forests have greater impacts on richness in the tropics. In other words, taller canopies generate biodiversity globally, but are more effective generators of biodiversity in the tropics.

# The mediating effect of verticality
The vertical layering of species within forests may mediate the capacity for tall forests to support rich assemblages. As forests increase in height, additional arboreal species may be added to the assemblage in novel canopy space [@macarthurBirdSpeciesDiversity1961; @rollLinkingVertebrateSpecies2015; @larueDiversityVolumeRelationships2023]. An increase in the number of arboreal species without an equivalent increase in terrestrial or fossorial species increases the verticality of the assemblage, which represents the skew in the vertical distribution of species toward the canopy [@oliveiraVerticalStratificationInfluences2019; @scheffersVerticalArborealityHorizontal2017]. Within wooded ecosystems, we therefore expect regions exhibiting high verticality to support species rich assemblages and steep increases in richness with canopy height [@oliveiraVerticalStratificationInfluences2019]. To quantify the verticality of an assemblage, we use the aforementioned trait databases, assigning verticality scores to each species based on their occurrence in fossorial (0), terrestrial (0.5), and arboreal (1) habitats [@oliveiraVerticalStratificationInfluences2019; @scheffersVerticalArborealityHorizontal2017]. For species occurring in multiple habitats, we averaged scores across habitats. We then calculated verticality as the standard effect size of the average verticality score across species in an assemblage, using 100 simulations to determine if assemblages are more vertical than expected by random chance given their richness and the evolutionary history of the biogeographic realm [@oliveiraVerticalStratificationInfluences2019] (Supplementary methods).

Verticality peaked in the tropics and declined toward the poles, exhibiting strong correlations with verticality across taxa (Fig. \@ref(fig:fig2)). Moist tropical forests contained assemblages comprised of x% to x% species exhibiting arboreal strategies (SUPP FIG), corresponding to some of the most species rich assemblages. Though temperate and boreal regions contain tall forests, albeit not quite reaching the heights of the tallest tropical canopies, they contain assemblages with much lower verticality, and higher proportions of species exhibiting terrestrial and fossorial strategies. This pattern suggests that the capacity for taller canopies to support higher richness declines when canopies cannot support highly vertical species assemblages. Understanding environmental factors that constrain the evolution of diverse arboreal assemblages is therefore key to unraveling the vertical dimension to Humbolt's enigma.

# Vegetation structure and canopy climates impact the capacity for tall canopies to support vertical communities
To examine the environmental factors influencing assemblage verticality, we used spatial linear mixed models to model assemblage verticality for each vertebrate class as a function of canopy height and its interaction with macroclimatic factors that may constrain or support arboreal life strategies, including minimum temperature of the coldest month, maximum temperature of the warmest month, precipitation of the driest month, and precipitation of the warmest month, as well as with macroclimatic factors that may influence the distinctness of vertical climate banding, including diurnal temperature range (Supplementary methods). We additionally included vegetation density in the models, which impacts the availability of habitat volume. While foliage height diversity and temperature seasonality also impact niche opportunities and the distinctness of climate banding, these variables where strongly correlated with canopy height and minimum temperatures, respectively, and were therefore excluded from the models. The macroclimate variables included in these models represent climate conditions experienced in open habitats rather than the microclimate conditions across forest profiles [@defrenneWeatherStationsLack2016; @bramerAdvancesMonitoringModelling2018]. While using macroclimate data was necessary due to the lack of global vertically resolved microclimate data, its use inhibits our ability to understand how microclimates experienced by forest-dwelling organisms impact community structure [@klingesProximalMicroclimateMoving2024; @zellwegerForestMicroclimateDynamics2020]. We therefore additionally modeled vertical microclimate profiles in six forests from tropical to boreal latitudes by integrating information on the vertical distribution of vegetation derived from NASA's GEDI spaceborne lidar mission [@dubayahGlobalEcosystemDynamics2020] with ERA5 macroclimate data at a high temporal resolution [@hersbachERA5HourlyData2023] in a process-based microclimate model that resolves within-canopy processes to predict vertical climate gradients [@macleanMicroclimfFastCanopy2024]. To reduce variability in vertical profiles not attributed to latitudinal differences, we restricted microclimate models to points below 500 m a.s.l and that represented the top 75th to 98th quantiles of canopy height within a 1° x 1° grid cell. By estimating the climate conditions experienced by forest-dwelling organisms, these models provide insight into the microclimatic processes that underlie macroclimate patterns and allow canopies to support vertically stratified and species rich assemblages.

Our models of verticality indicate that taller canopies support highly vertical vertebrate assemblages, while dry, and thermally variable macroclimate conditions, including extreme hot and cold temperatures limit verticality and mediate the capacity for tall forests to support vertical assemblages, albeit to varying degrees across vertebrate classes (Fig. \@ref(fig:fig3)). These patterns arise due to impacts of spatial climate variation on vegetation structure and physiological constraints to arboreal lifestyles as well as impacts of temporal climate variation on the distinctness of vertical climate bands.

## Climate produces structural constraints to verticality
Climate impacts the structural complexity of vegetation, which plays a key role in the the availability of arboreal niche opportunities and is typically associated with high animal richness [@xingEcologicalPatternsProcesses2023a; @daviesAdvancesAnimalEcology2014]. High dry season precipitation and thus low precipitation seasonality may support highly vertical assemblages by enabling development of tall and structurally complex forests with densely layered vegetation and generally high tree diversity [@ehbrechtGlobalPatternsClimatic2021a; @decontoCharacterizingStructuralComplexity2024; @hakkenbergClimateMediatesRelationship2021]. Wet climates also support diverse epiphyte communities, which provide additional microhabitats in the canopy that contribute to high arboreal richness [@taylorVascularEpiphytesContribute2022; @thomsenSecondaryFoundationSpecies2018; @ellwoodDoublingEstimateInvertebrate2004]. Hotspots of structural complexity that coincide with some of the tallest forests, including the Choco-Darien, the Amazon, Borneo, and New Guinea [@ehbrechtGlobalPatternsClimatic2021a; @decontoCharacterizingStructuralComplexity2024], thus provide opportunities for fine-scale niche partitioning, yielding species rich assemblages [@camargoForestVerticalComplexity2018; @gouveiaForestStructureDrives2014; @oliveiraVerticalStratificationInfluences2019]. In contrast, lower structural complexity in parts the Afrotropics [@ehbrechtGlobalPatternsClimatic2021a] may reduce niche opportunities for arboreal species, yielding less vertical assemblages and lower richness than expected based on the height of the canopy (Fig. \@ref(fig:fig1)).

At higher latitudes, lower precipitation and light hinder the development of densely layered forests by restricting the functional and structural configurations of trees [@ehbrechtGlobalPatternsClimatic2021a; @hakkenbergClimateMediatesRelationship2021; @atkinsPowerLawScaling2022]. In evergreen needleleaf forests that dominate boreal regions, these environmental constraints dampen the rate of increase in structural complexity with canopy height [@atkinsPowerLawScaling2022]. Tall canopies therefore provide fewer niche opportunities in boreal regions, which may constrain assemblage verticality and limit species richness relative to canopies of equivalent height at lower latitudes. Yet, hotspots of structural complexity in temperate forests, including North Pacific Alaskan, Valdivian, and Tasmanian forests, that nearly match the complexity of tropical forests [@ehbrechtGlobalPatternsClimatic2021a] also contain assemblages with lower verticality than the tropics and fewer species than expected based upon their canopy height (Figs. \@ref(fig:fig1), \@ref(fig:fig2)). MacArthur & MacArthur [-@macarthurBirdSpeciesDiversity1961] observed this pattern over 50 years ago, noting that "tropical bird species diversity is much greater than the temperate one for habitats of comparable profiles." The interaction between macroclimate and vertical microclimate profiles provides an explanation that mirrors the climatic mechanisms allowing mountains to be better generators of biodiversity in the tropics [@rahbekHumboldtsEnigmaWhat2019].

## The influence of canopy climates

Vertical features in the landscape, including mountains and trees bring distinct climates within close proximity of each other [@rahbekHumboldtsEnigmaWhat2019; @leahyVerticalNicheElevation2021; @vinodThermalSensitivityForest2023]. By absorbing and reflecting solar radiation and reducing vertical air mixing, dense forests produce vertical microclimate gradients that capture a remarkable volume of climatic variation given their geographic footprint [@defrenneForestMicroclimatesClimate2021]. These steep climate gradients can support high rates of turnover from the ground to the canopy that yield vertically stratified and species rich assemblages [@bashamVerticalStratificationCollapses2020; @xingSpeciesTurnoverAnt2022; @oliveiraVerticalStratificationInfluences2019; @camargoForestVerticalComplexity2018]. However, more extreme microclimates in the canopy than the understory may place physiological constraints on arboreal lifestyles depending on background macroclimate conditions, effectively truncating the available volume of forest and limiting the capacity for tall forests to support high richness, just as lower elevation snow lines limit the capacity for tall mountains to generate biodiversity at higher latitudes [@oliveiraVerticalStratificationInfluences2019].

For example, our models indicate that low dry season precipitation constrains verticality across vertebrate classes (Fig. \@ref(fig:fig3)). These macroclimate patterns were accompanied by steep gradients in minimum relative humidity, which declined x% from the ground to the top of the canopy (Fig. \@ref(fig:fig4)). Low precipitation may exaggerate the drier canopy conditions and risk desiccation in hot climates, where arboreal species depend on evapotranspiration to maintain body temperatures below critical thermal maxima, or in cold climates, where basking to increase body temperature can quickly lead to desiccation without adequate moisture availability or physiological defenses. By imposing environmental filters on the functional morphologies that can inhabit canopy environments, dry climates therefore reduce assemblage verticality and limit the number of species that an increase in canopy height can support, in line with the physiological tolerance hypothesis [https://onlinelibrary.wiley.com/doi/full/10.1111/j.1461-0248.2004.00671.x]. This pattern is evident for frog communities in the seasonally dry forests of Panama, where collapse of vertical stratification toward the ground during the dry season due to water limitation in the canopy is accompanied by richness declines and homogenization of arboreal diversity [@bashamVerticalStratificationCollapses2020]. Simply put, extreme canopy microclimates that physiologicall constrain arboreal life truncate the volume of available habitat, which prevents tall trees from supporting species rich assemblages. In contrast, high dry season precipitation in tropical moist forests supports highly vertical assemblages by reducing physiological constraints on arboreal life strategies in the upper canopy. Species are therefore able to finely partition the entire volume of forested habitat, leading to a dense packing of species in the canopy and allowing increases in canopy height to yield rapid increases in species richness (Fig. \@ref(fig:fig1)).

Hot and cold temperatures had less consistent effects across taxa, but may still play critical roles in physiological constraints, particularly in regions exhibiting steeper vertical gradients characterized by hotter maximum and colder minimum temperatures. For example, high maximum temperatures reduced verticality and/or [REDUCED THE POSITIVE IMPACT OF CANOPY HEIGHT ON VERTICALITY FOR REPTILES, MAMMALS, AND BIRDS] across vertebrate classes, while low minimum temperatures reduced positive effects of canopy height for birds and reptiles (Fig. \@ref(fig:fig3)). High maximum temperatures may cause ectotherm body temperatures to exceed critical thermal maxima (McKechnie et al. 2017; Sunday et al. 2014), particularly in drier regions where maximum temperatures peaked. Microclimate models additionally indicated that vertical maximum temperature gradients were steepest in the temperate forests, though canopy temperatures exceeding 40C may limit the capacity for species to stratify across the entire vertical climate gradient (Fig. \@ref(fig:fig4)). Similarly, low minimum temperatures can cause body temperatures to exceed critical thermal minima for ectotherms or impose energetic constraints to endotherms, particularly in the windier conditions that characterize the canopy  (Sunday et al. 2014) (Byman, Hay, and Bakken 1988). Shallow vertical minimum temperature gradients (Fig. \@ref(fig:fig4)), particularly during leaf off conditions [CITE], cause environmental filters to act across the entire canopy, posing severe limitations on the capacity for taller canopies to increase species richness.

[https://nsojournals.onlinelibrary.wiley.com/doi/full/10.1111/ecog.07122]

Temporal climate variability also plays a key role in the capacity for verticality to mediate biogeographic patterns by impacting climatic distinctness. [relationships between species richness and canopy height by impacting climatic distinctness]. In montane regions, climatic distinctness across elevations declines with latitude due to increasing seasonal temperature variation toward the poles [@rahbekHumboldtsEnigmaWhat2019; @janzenWhyMountainPasses1967]. Low seasonal variation in the tropics produces narrow and distinct climate bands that exhibit little overlap across elevations, yielding high rates of turnover that allow tropical mountains to generate extraordinary levels of biodiversity [@rahbekHumboldtsEnigmaWhat2019]. In contrast, high thermal variation in temperate regions increases the similarity between climates experienced at high and low elevations, reducing the capacity for temperate mountains to support high rates of turnover and high levels of biodiversity [@rahbekHumboldtsEnigmaWhat2019].

Our models suggest that a similar mechanism may contribute to the capacity for tropical canopies to support assemblages with higher verticality and species richness than temperate canopies of equivalent heights. At the macroclimate scale, models of verticality generally indicated no significant impact of diurnal temperature variation, though low diurnal temperature variation increased verticality of mammals (Fig. \@ref(fig:fig3)). However, microclimate models indicated clear increases in daily temperature variation in higher latitude forests (Fig. \@ref(fig:fig4)). Moreover, the increase in thermal variability from the ground to the canopy was steeper in temperate and boreal regions. Temperate and boreal canopies canopies therefore exhibited less distinct climate bands, with vertical zonation collapsing nearly completely during morning hours. Without unique climatic niches across the forest profile, these regions are unlikely to support rapid vertical turnover, limiting the number of species that may be added to the community given an increase in canopy height even in regions with climate conditions that physiologically support arboreal life. In contrast, tropical canopies maintained distinct climate bands throughout the day, with the lowest and highest bands exhibiting little to no temperature overlap. Narrow and distinct vertical climate niches may therefore support rapid vertical turnover and increases in richness with canopy height. that contributes the capacity for tropical  unique communities

- climate variation increases vertical range size of species - Leahy

# Conclusion - bring back to Humbolt's engima

Yet only recently has the three-dimensional nature of terrestrial landscapes been recognized as a critical dimension to our understanding of biogeographic patterns [@fredstonReimaginingSpeciesMove2025; @xingEcologicalPatternsProcesses2023a].

Since verticality research was started, most research has been on impacts of vegetation structure. Yet no hypotheses had been proposed regarding inconsistent conclusions of the impacts of vegetation structure on richness.

The canopy supports fewer arboreal species at high latitudes because
1) physiological limitations and physiological tolerance hypothesis
2) high temporal climate variabiltiy in high latitude canopies reduces vertical zonation, limiting vertical turnover and therefore richness

- Trees, just as mountains, bring different climate zones within close proximity of each [@rahbekHumboldtsEnigmaWhat2019]. While the climatic difference from the ground to the canopy is much less than the climatic difference across thousands of meters in elevation, the climatic rate of change is often faster from the ground to the canopy. This climatic layering allows distinct assemblages of species to exist across forest strata.

- climate stability, range size, and niche width - greater climate stability in the tropics allows for more specialization within the canopy, while higher climate variability in the canopy in high latitudes reduces specialization and turnover. Compare to the ground - see Dave's mountaintops paper?

- The capacity for tall forests to support rich arboreal assemblages depends on the density of vegetation across the vertical profile and the presence of canopy microclimates that limit physiological constraints on arboreal life strategies.

- In high latitude forests, tall canopies may help generate biodiversity on the ground by buffering microclimate and providing small increases in structural complexity that can be inhabited by few arboreal animals with specialized physiologies. However, harsh microclimates in the upper canopy largely truncate niche availability, preventing fine-scale partitioning of canopy habitats by arboreal species and thus limiting the extent to which taller canopies increase species richness.

- The biodiversity-structure hypothesis suggests xyz...
  - Greater veg density across vertical profile = higher productivity

[INTEGRATE VERTICAL COMPONENT INTO THIS PARAGRAPH]

- comparison to mountains
  - climate bands are narrower in the tropics -> higher turnover rates
  - we don't see this pattern for vertical climate profiles (similar shapes in profiles across latitudes and less variation in the troipcs)
- results
  - the vertical difference in rhmin and tmax is greater than the latitudinal difference from tropical to boreal forests
  - stable microclimates in tropical canopies -> specialization and speciation

In climate conditions that do not constrain verticality, shifts up the canopy are associated with radiations within the canopy that produce positive relationships between canopy verticality and species richness
In climate conditions that constrain verticality, arboreal species occur, but arboreal radiations do not, causing neutral or negative relationships between verticality and richness (although this mechanism/relationship is noisy)

# Methods
## Data
The canopy height layer was obtained from Lang et al. [-@langHighresolutionCanopyHeight2023], which represents canopy height in the year 2020 at a 10 m resolution. To obtain canopy height in forests, we aggregated the data to match the resolution of Hansen et al. [-@hansenGlobalLandUse2022] global land cover and land use in 2019 (~ 30 m resolution) and used this layer to mask all grid cells with woody vegetation < 3 m, surface water and permanent ice, built-up land, cropland, recent forest loss, and land use associated with forestry and shifting cultivation (e.g., oil palm). We then aggregated and averaged the remaining pixels to the resolution of the verticality data (50 km).

We classified biomes into larger categories based on the 2017 biome categories. For biomes that occurred across latitudes (e.g., Montane Grasslands and Shrublands), we classified them as tropical if they occurred within the latitudes covered by biomes classified as tropical and temperate if they occurred within the latitudes classified as temperate. Additionally, Mediterranean forests, woodlands, and scrub were partitioned into temperate forests if the ecoregion contained "forest" and temperate woodland, savanna, and shrubland if the ecoregion name did not contain the word "forest".

MASS was used to produce models of canopy height by richness. sdmTMB was used to produce models of verticality. DHARMa was used to check model fit and assumptions. ggeffects was used to generate model predictions and confidence intervals. Figures were generating using tidyterra, patchwork, ggplot2

# Data availability statement
canopy height:

# Figures

```{r fig1, fig.id="fig1"}
# Fig. 1

# Legend: Spatial distribution of a) canopy height, b) vertebrate richness (all vertebrates combined), and
# c) deviance residuals for a negative binomial model of total vertebrate richness by canopy height
# d) Relationships between species richness and canopy height for each taxon in each
# biome category. Dots indicate observed values and lines indicate negative binomial
# model fits of richness by canopy height.

"figures/ms_figures/fig1_canopy_richness.png"

```

```{r fig3, fig.id="fig3", fig.cap="Coefficients of standardized predictors for spatial LMMs of verticality as a function of environmental factors. Models were conducted for each vertebrate class. Error bars represent 95% confidence intervals. Filled circles with solid error bars have confidence intervals that do not overlap zero. Open circles with dashed error bars have confidence intervals that overlap zero. Tmin~cold~ = minimum temperature of the coldest month; Tmax~warm~ = maximum temperature of the warmest month; Temp~diu~ = diurnal temperature range; Precip~dry~ = precipitation of the driest quarter; Precip~warm~ = precipitation of the warmest quarter.", fig.width=140, fig.height=120}

# Fig. 3

"figures/main_figs/fig3_sdmTMB_coefs.png"

```

THINGS TO FIX IN FIG4
- diurnal temp range is for warmest month in strata plot and coldest month in profile plot - make consistent
- replace map with concept figure similar to Rahbek
- should we add strata plots for other variables? Or add to supplement - probable supplement
- text sizing
- point size in map if we keep the map
```{r fig4, fig.id="fig4", fig.cap=", fig.width=140, fig.height=120}

# Fig. 4

"figures/ms_figures/fig4_profiles.png"

```
# Supplementary figures

```{r splot-global-chrich, fig.id="global-chrich", width=90, height=90, fig.cap="Global relationship between total vertebrate richness and canopy height. Line indicates predictions from a negative binomial model of richness by canopy height ± 95% CI. Points indicate observed richness by canopy height.}

"figures/ms_figures/supp_figs/total_richness~canopy.png"
```

```{r splot-resids-taxa, fig.id="resids-taxa", width=180, height=100, fig.cap="Deviance residuals from negative binomial models of richness by canopy height for each vertebrate class: a) birds, b) mammals, c) reptiles, d) amphibians."}

"figures/ms_figures/supp_figs/deviance_resids_taxa.png"
```

# Supplementary material

## Microclimate models

We selected 6 1 x 1 degree grid cells, with 2 each in tropical, temperate, and boreal forests. We downloaded GEDI points that intersected with these cells. All GEDI points were leaf on conditions due to low accuracy of vertical profiles during leaf-off conditions. When modeling climate in the coldest month, this resulted in only few points in temperate regions, as these were restricted to mixed forests and did not include deciduous forests.

# SCRATCH SPACE
  - TRY RELATIVIZING RATES OF INCREASE BECAUSE THERE ARE MORE SPECIES IN THE TROPICS. EG. INCREASE IN 1 SPECIES OUT OF 10 TOTAL SPECIES = INCREASE IN 10 SPECIES IF THERE ARE 100 SPECIES IN THE REGION
 - on second thought, I don't think this is necessary because we want to know why more species can't exist in trees in boreal regions

Just as topographic complexity and climate explain Humbolt's enigma in mountains, vegetative complexity and climate may play key roles in mediating the relationship between canopy height and species richness by impacting the availability of vertical niche opportunities and physiologically constraining arboreal lifestyles [CITE]. Densely layered forest structures not only provide habitat across the forest profile, but produce vertical microclimate gradients via interactions with macroclimate conditions, solar radiation, and wind speed. From the humid and thermally stable understory to the dry and thermally variable canopy that experiences both hot and cold temperature extremes [CITE], this vertical gradient may interact with latitudinal climate gradients to help or hinder the capacity for tall forests to generate high levels of biodiversity. For example, increased exposure to extreme cold in the canopies of high latitude forests may truncate the volume of habitat that can support arboreal species and thus limit richness despite tall canopies [@oliveiraVerticalStratificationInfluences2019]. Moreover, relatively low thermal variability in the canopies of tropical forests may increase the distinctness of vertical climate bands that support high turnover and thus enhance the capacity for tall forests to generate biodiversity in the tropics [CITE].

in wet tropical forests, macroclimate conditions produce warmer and wetter canopy microclimates, which reduce thermal and hydric stress and may allow species to finely partition the entire three-dimensional habitat space [@oliveiraVerticalStratificationInfluences2019]. Taller canopies in tropical regions could thus support higher animal verticality and richness. In contrast, extreme cold, or the combination of heat and drought may impede the evolution of arboreal lifestyles, constraining biodiversity to lower vertical positions tat are more climatically buffered (cite). Harsh macroclimate conditions may therefore truncate the availability of vertical habitat space, thereby limiting vertical stratification and species richness despite the presence of vegetation.

Integrating new technologies can shed light on the mechanisms driving decades-old ecological patterns.

But only within the last decade have we been able to measure forest structure and acquire data on vertebrate use of canopy strata at a global extent, allowing us to now gain a comprehensive understanding of the impact of vertical habitat structure on species richness patterns.

In forests, species assemblages stratify vertically along resource and microclimate gradients that extend from beneath the ground to the top of the canopy [@bashamVerticalStratificationPatterns2023a; @xingEcologicalPatternsProcesses2023a]. Tall, multi-layered forests increase both habitat volume and niche opportunities, enabling species to partition habitat across vertical strata akin to multi-floored apartments [@macarthurLimitingSimilarityConvergence1967; @macarthurBirdSpeciesDiversity1961; @larueDiversityVolumeRelationships2023].

where verticality declines, species may use terrestrial habitats during the dry season to maintain moisture requirements assemblage verticality declines because species exhibit semi-arboreal life strategies, which reduces assemblage verticality.

"short tropical forests show higher complexity relative to other forests of similar heights" - de Conto et al. 2023

This document presents most of the features of the package `r ftext("officedown", ft)`. 
`r fp`

## Table of content

<!---BLOCK_TOC--->

## List of figures

<!---BLOCK_TOC{seq_id: 'fig'}--->

## List of tables

<!---BLOCK_TOC{seq_id: 'tab'}--->

\newpage

## Sections

### A two columns section

<!---BLOCK_MULTICOL_START--->

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna. Sed facilisis dapibus diam nec volutpat. Maecenas facilisis dapibus egestas. Curabitur dignissim pharetra pulvinar. Nunc bibendum elit sed cursus congue.

`r run_columnbreak()`Curabitur ligula quam, iaculis faucibus orci quis, vestibulum lobortis lectus. Suspendisse fringilla nisl pulvinar, laoreet tellus sed, sollicitudin tortor. Donec consequat congue erat in iaculis. Curabitur luctus tellus ut turpis iaculis, nec laoreet ligula scelerisque.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna. Sed facilisis dapibus diam nec volutpat. Maecenas facilisis dapibus egestas. Curabitur dignissim pharetra pulvinar. Nunc bibendum elit sed cursus congue. Curabitur ligula quam, iaculis faucibus orci quis, vestibulum lobortis lectus. Suspendisse fringilla nisl pulvinar, laoreet tellus sed, sollicitudin tortor. Donec consequat congue erat in iaculis. Curabitur luctus tellus ut turpis iaculis, nec laoreet ligula scelerisque.

<!---BLOCK_MULTICOL_STOP{widths: [3,3], space: 0.2, sep: true}--->

### End of two columns section

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna. Sed facilisis dapibus diam nec volutpat. Maecenas facilisis dapibus egestas. Curabitur dignissim pharetra pulvinar. Nunc bibendum elit sed cursus congue. Curabitur ligula quam, iaculis faucibus orci quis, vestibulum lobortis lectus. Suspendisse fringilla nisl pulvinar, laoreet tellus sed, sollicitudin tortor. Donec consequat congue erat in iaculis. Curabitur luctus tellus ut turpis iaculis, nec laoreet ligula scelerisque.

<!---BLOCK_LANDSCAPE_START--->

### Landscape section

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna. Sed facilisis dapibus diam nec volutpat.

```{r, echo=FALSE}
barplot(1:5, col=1:5)
```

<!---BLOCK_LANDSCAPE_STOP--->

### Normal section

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna. Sed facilisis dapibus diam nec volutpat. Maecenas facilisis dapibus egestas. Curabitur dignissim pharetra pulvinar. Nunc bibendum elit sed cursus congue. Curabitur ligula quam, iaculis faucibus orci quis, vestibulum lobortis lectus. Suspendisse fringilla nisl pulvinar, laoreet tellus sed, sollicitudin tortor. Donec consequat congue erat in iaculis. Curabitur luctus tellus ut turpis iaculis, nec laoreet ligula scelerisque.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nulla eu pulvinar arcu, quis aliquam dui. In at cursus ante. Vestibulum non sagittis lacus. Duis vitae iaculis dui. Vivamus tempor, nibh ut pretium tempus, enim lorem dignissim quam, at euismod massa magna at magna. Sed facilisis dapibus diam nec volutpat. Maecenas facilisis dapibus egestas. Curabitur dignissim pharetra pulvinar. Nunc bibendum elit sed cursus congue. Curabitur ligula quam, iaculis faucibus orci quis, vestibulum lobortis lectus. Suspendisse fringilla nisl pulvinar, laoreet tellus sed, sollicitudin tortor. Donec consequat congue erat in iaculis. Curabitur luctus tellus ut turpis iaculis, nec laoreet ligula scelerisque.

## Tables

### Table 1

```{r tab.cap="caption 1", tab.id="mtcars"}
head(mtcars)
```

### Table 2

```{r tab.cap="iris"}
head(iris)
```

### Table 3

```{r tab.cap="cars", tab.id="cars"}
head(cars)
```

## figures

### A boxplot

```{r fig.cap="A boxplot", fig.id = "boxplot"}
boxplot(1:8)
```

### A barplot

```{r fig.cap="What a barplot", fig.id = "barplot"}
barplot(1:8)
```

## Lists

Amet nunc eros curabitur tellus massa, eros maximus porttitor sociosqu, pellentesque.

* Erat mauris egestas finibus tincidunt sed in rhoncus a tellus etiam. 
    - A adipiscing per ultricies justo tellus lorem. 
        - Imperdiet ut dui primis, sed gravida, at sed nulla. 
        - Sem posuere lacus consequat inceptos dapibus duis malesuada finibus. 
    - Urna sed dui, ornare, eu turpis mus pellentesque amet amet bibendum. 
* Himenaeos tincidunt, auctor dapibus scelerisque, montes nunc faucibus 
sodales malesuada ridiculus sed cubilia ligula.

1. Erat mauris egestas finibus tincidunt sed in rhoncus a tellus etiam. 
    1. A adipiscing per ultricies justo tellus lorem. 
        1. Imperdiet ut dui primis, sed gravida, at sed nulla. 
        2. Sem posuere lacus consequat inceptos dapibus duis malesuada finibus. 
    2. Urna sed dui, ornare, eu turpis mus pellentesque amet amet bibendum. 
2. Himenaeos tincidunt, auctor dapibus scelerisque, montes nunc faucibus 
sodales malesuada ridiculus sed cubilia ligula.

Lorem dolor interdum orci eros pellentesque semper tristique, sodales, et sed
ut! Porta mattis natoque et. Ac facilisi ipsum viverra elementum vestibulum
ligula amet enim magnis luctus ullamcorper. Rhoncus rhoncus elit in at nisl.
Tincidunt habitant sit.

* Aptent conubia quam montes id sagittis.
    - Mattis nisi nascetur, aliquam duis ex, tristique.
        - Imperdiet ut dui primis, sed gravida, at sed nulla. 
        - Donec ligula nulla ac. Nisl ac at accumsan sagittis eros felis lobortis amet 
        nec phasellus urna bibendum sapien.
    - Eu dui ac id, dictum proin consectetur convallis.
* Facilisi eu lectus mauris lorem. Et sed sapien pellentesque sed etiam vehicula. 
* In porttitor id lorem eu efficitur, nisl dis!

## Reference

see figure \@ref(fig:boxplot) and table \@ref(tab:mtcars)!