### 2025-07-28 manuscript.Rmd
adjust figure 1 to show all taxa for scatter plot

---
bibliography: "r-refs.bib"
csl: science.csl
output: 
  officedown::rdocx_document:
    reference_docx: "style-guide.docx"
    mapstyles:
      Normal: ['First Paragraph']
plots:
  caption:
    style: Image Caption
    pre: 'Figure '
    sep: ': '
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \rhead{Global Verticality}
- \begin{document}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE, warning=FALSE, message=FALSE, error=FALSE)
library(officedown)
library(officer)
library(tidyverse)
library(spatialreg)
library(data.table)
library(flextable)
library(knitr)
library(chunkhooks)

fp <- fp_par(
  text.align = "center", 
  padding.bottom = 20, padding.top = 120, 
  border.bottom = fp_border())

ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)

hook_figure_unit("mm") # sets figure units to mm
```

```{r}
# analysis scripts
# storing all figures in figures/ms_figures/
# other figures are from older versions
```

::: {custom-style="Author"}

**Forest canopies are better generators of biodiversity in the tropics**

Lydia G. Soifer^1\*^, David P. Edwards^2,3^, Brunno F Oliveira^4,5^, Mario R Moura^6^, Brett R. Scheffers^4^

1 School of Natural Resource and Environment, University of Florida, USA

2 Department of Plant Sciences and Centre for Global Wood Security, University of Cambridge, Cambridge, UK

3 Conservation Research Institute, University of Cambridge, Cambridge, UK

4 Centre de Synthèse et d’Analyse sur la Biodiversité, Montpellier, France

5 Department of Wildlife Ecology and Conservation, University of Florida, USA

6 Departamento de Sistemática e Ecologia, Universidade Federal da Paraíba, João Pessoa, Brazil

Email addresses:

^\*^ lysoifer@gmail.com (corresponding author)

dpe29@cam.ac.uk

brunno.oliveira@me.com

mariormoura@gmail.com

brett.scheffers@ufl.edu

:::

\newpage

# Abstract

::: {custom-style="Normal"}

Three-dimensional landscape structure, characterized by mountains and trees, plays a key role in shaping species richness patterns. Taller forests have been proposed to support high levels of biodiversity by increasing habitat volume and producing vertical zones of microclimate and resource availability that allow species to partition niche opportunities from the ground to the canopy. Yet the capacity for taller forests to support greater biodiversity is inconsistent globally. Here, we examine this pattern and the climatic mechanisms from macro to micro scales may mediate the capacity for tall canopies to increase species richness via impacts on the vertical structure of vertebrate communities. Overall, we find that tall canopies support greater biodiversity globally, but high precipitation allows them to yield more rapid increases in richness in the wet tropics by supporting structurally complex vegetation and reducing physiological constraints imposed by dry canopy microclimates.

:::

\newpage

::: {custom-style="Normal"}

For centuries, ecologists have been intrigued by the role of vertical landscape features in biogeographic patterns [@vonhumboltEssayGeographyPlants2013; @macarthurBirdSpeciesDiversity1961; @bashamVerticalStratificationPatterns2023a; @xingEcologicalPatternsProcesses2023a]. In 1799, Alexander von Humboldt began an expedition that would ultimately canonize the role of mountains as generators of species richness, particularly in tropical regions [@vonhumboltEssayGeographyPlants2013; @HumboldtsLegacy2019; @rahbekHumboldtsEnigmaWhat2019]. This pattern—'Humboldt's enigma'—can be attributed to topographic and climatic variation [@rahbekHumboldtsEnigmaWhat2019]. The vertical nature of mountains provides more land surface per unit area than flat regions and brings distinct climates within close proximity of each, which produces rapid rates of species turnover. Yet mountains are better generators of biodiversity in the tropics because high latitude mountains have lower elevation snow lines that truncate the area of land suitable for diverse communities, and greater seasonal temperature variation, which reduces the climatic distinctness of elevational bands relative to tropical mountains [@rahbekHumboldtsEnigmaWhat2019; @vonhumboltEssayGeographyPlants2013; @janzenWhyMountainPasses1967].

Over a century later, Robert MacArthur highlighted the role of another vertical dimension in promoting high biodiversity: the vertical structure of vegetation [@macarthurBirdSpeciesDiversity1961]. Tall, multi-layered forests generate biodiversity through similar mechanisms to mountains. By increasing habitat volume and producing vertical zones of microclimate and resource availability, tall forests allow species to vertically stratify habitat akin to multi-floored apartments [@bashamVerticalStratificationPatterns2023a; @xingEcologicalPatternsProcesses2023a; @macarthurLimitingSimilarityConvergence1967; @macarthurBirdSpeciesDiversity1961; @larueDiversityVolumeRelationships2023]. This vertical stacking of species with increasing arboreal niche opportunities has been described as a mechanistic link between canopy height and species richness, and is thought to contribute to exceptionally high biodiversity in the tropics where canopies are often tall and structurally complex [@oliveiraVerticalStratificationInfluences2019; @rollLinkingVertebrateSpecies2015; @fengForestCanopyHeight2020; @decontoCharacterizingStructuralComplexity2024]. Yet taller canopies are not consistently associated with higher biodiversity across the globe [@larueDiversityVolumeRelationships2023; @macarthurBirdSpeciesDiversity1961; @rollLinkingVertebrateSpecies2015; @fengForestCanopyHeight2020]. The capacity for forests to generate biodiversity via vertical stratification, yet the variation in their capacity to do so globally therefore presents 'MacArthur's enigma'.

Despite decades of ecological inquiry into the vertical dimensions of habitat structure and biodiversity [@macarthurBirdSpeciesDiversity1961; @scheffersIncreasingArborealityAltitude2013; @xingEcologicalPatternsProcesses2023a; @rollLinkingVertebrateSpecies2015], the mechanisms that help or hinder the capacity for tall forests to produce species rich assemblages remain poorly understood yet are critically important for understanding biogeographic patterns, impacts of climate and land-use change on biodiversity, and the conservation value of forests. Following explanations of Humboldt's enigma [@rahbekHumboldtsEnigmaWhat2019], we examine how climate from macro to micro scales mediates the relationship between canopy height and species richness by impacting the capacity for tall canopies to support vertically stratified and species rich arboreal assemblages. For example, climatic conditions that promote structurally complex canopies may provide more niche opportunities, allowing tall canopies to support more biodiversity than structurally simpler canopies of equivalent height [@daviesAdvancesAnimalEcology2014; @oliveiraVerticalStratificationInfluences2019]. Moreover, physiological constraints imposed on arboreal life strategies by dry, hot, or cold canopy microclimates combined with the distinctness of vertical climate bands may impact species turnover from the ground to canopy and thus the capacity for taller canopies to support increasing richness [@bashamVerticalStratificationCollapses2020; @oliveiraVerticalStratificationInfluences2019]. Only recently have we been able to address these hypotheses, as our ability to quantify global variation in 3D vegetation structure, microclimates, and animal community structure has increased with the launch of satellite missions to measure forest height and structural complexity [@popkinSpaceLaserMap2018], the development of microclimate models that can estimate 3D microclimates anywhere on Earth [@macleanMicroclimfFastCanopy2024], and the collation of large ecological databases on vertical habitat use [@mouraPhylogenyinformedCharacterisationGlobal2024a; @wilmanEltonTraitsSpecieslevelForaging2014].

We combine range maps of over 24,000 terrestrial vertebrate species with comprehensive data on vertical habitat use from the TetrapodTraits and EltonTraits databases [@mouraPhylogenyinformedCharacterisationGlobal2024a; @wilmanEltonTraitsSpecieslevelForaging2014] to examine how climate from macro to micro scales may mediate the capacity for tall forests to support diverse assemblages via impacts on the vertical structure of vertebrate communities. We first investigate the capacity for taller canopies to support increasing species richness in woody biomes across latitudes. We then examine the processes driving observed patterns using a combination of global models of vertical assemblage structure parameterized with macroclimate and forest structure data and regional models of vertical microclimate profiles [@macleanMicroclimfFastCanopy2024; @macleanMicropointCanopyPoint2025], which integrate forest structure data from NASA's Global Ecosystem Dynamics Investigation (GEDI) [@dubayahGlobalEcosystemDynamics2020] with high temporal resolution macroclimate data [@hersbachERA5HourlyData2023] to estimate temperature and relative humidity from the ground to the canopy.

:::

# Tall canopies generate biodiversity, but more so in the tropics

```{r}
source("scripts/03_analyses/01_canopyheight_richness_relationships.R")
# ratedifs provides table of differences in richness between forests that are 15 and 25 m tall for each biome/taxa combo
```

::: {custom-style="Normal"}

To evaluate the capacity for tall canopies to generate biodiversity, we fit negative binomial regression models of global vertebrate richness as a function of canopy height for all vertebrates combined and each vertebrate class (birds, mammals, reptiles, and amphibians), mapping the deviance residuals to determine where canopy height over- or under-predicted species richness. Globally, taller canopies support more species (Fig. S\@ref(fig:global-chrich); Table S\@ref(tab:richness-mod-results)), because they increase habitat volume and complexity, which allows fine-scale niche partitioning from the ground to the canopy [@larueDiversityVolumeRelationships2023; @macarthurBirdSpeciesDiversity1961; @xingEcologicalPatternsProcesses2023a]. Yet global models of vertebrate richness by canopy height generally overpredict richness at high latitudes (negative residuals) and underpredict richness in the tropics (positive residuals) (Figs. \@ref(fig:fig1)c, S\@ref(fig:resids-taxa)). Therefore, canopies of equivalent height generally harbor more species in the tropics than in temperate and boreal regions.

Models that allow impacts of canopy height to vary across biomes indicate that increasing canopy height yields greater gains in richness in the wet tropics than at higher latitudes or drier tropical biomes (Fig. \@ref(fig:fig1)d). For example, an increase in canopy height from 15 m to 25 m corresponds to an increase of `r round(ratedifs$dif[ratedifs$biome2 == "Tropical forest" & ratedifs$taxa=="Birds"])` bird and `r round(ratedifs$dif[ratedifs$biome2 == "Tropical forest" & ratedifs$taxa=="Amphibians"])` amphibian species in tropical forests, but only `r round(ratedifs$dif[ratedifs$biome2 == "Boreal" & ratedifs$taxa=="Birds"])` bird and `r round(ratedifs$dif[ratedifs$biome2 == "Boreal" & ratedifs$taxa=="Amphibians"])` amphibian species in boreal regions. The difference in impacts of taller canopies on biodiversity between wet and dry tropical regions is particularly evident in Madagascar, where the global model underpredicts vertebrate richness in the humid forests of the east coast and overpredicts richness in the drier woodlands and shrublands of the southwest (Figs. \@ref(fig:fig1)c, S\@ref(fig:resids-taxa)). Thus, increases in canopy height have greater positive impacts on richness in the wet tropics and lower impacts in drier regions or higher latitudes. Simply put, taller canopies generate biodiversity globally, but are more effective generators of biodiversity in the wet tropics - a pattern mirroring that of mountains [@rahbekHumboldtsEnigmaWhat2019].

:::

# The mediating effect of verticality

```{r}
source("scripts/03_analyses/02_vertmaps_and_vertRichness_correlations.R")

# mean proportion of arboreal species
parb.wettropics = round(as.numeric(parb.mean[parb.mean$biome=="Tropical & Subtropical Moist Broadleaf Forests", 2]),2)

```

::: {custom-style="Normal"}

Verticality—the skew in the vertical distribution of species toward the canopy [@oliveiraVerticalStratificationInfluences2019; @scheffersVerticalArborealityHorizontal2017]—peaked in the wet tropics and declined toward the poles, exhibiting strong correlations with richness across taxa (Fig. \@ref(fig:fig2), S\@ref(fig:meanvert-biome)). Tropical and subtropical moist forests contained assemblages with an average of `r parb.wettropics*100`% of species exhibiting arboreal strategies, corresponding to some of the most species rich regions (Fig. S\@ref(fig:meanprop-biome)). Though temperate and boreal regions contain tall forests, albeit not quite reaching the heights of the tallest tropical canopies, they contain assemblages with much lower verticality, and higher proportions of species exhibiting terrestrial and fossorial strategies (Fig. \@ref(fig:fig2)). This pattern suggests that the capacity for taller canopies to support higher richness declines when canopies cannot support highly vertical species assemblages. Understanding environmental factors that constrain the evolution of diverse arboreal assemblages is therefore key to unraveling MacArthur's enigma.

The vertical layering of species within forests may mediate the capacity for tall forests to support diverse assemblages [@oliveiraVerticalStratificationInfluences2019; @macarthurBirdSpeciesDiversity1961; @xingEcologicalPatternsProcesses2023a]. As forests increase in height, additional arboreal species may be added to the assemblage in novel canopy space [@macarthurBirdSpeciesDiversity1961; @rollLinkingVertebrateSpecies2015; @larueDiversityVolumeRelationships2023]. An increase in the number of arboreal species without an equivalent increase in terrestrial or fossorial species increases the verticality of the assemblage, which represents the skew in the vertical distribution of species toward the canopy [@oliveiraVerticalStratificationInfluences2019; @scheffersVerticalArborealityHorizontal2017]. Within wooded ecosystems, we therefore expect regions exhibiting high verticality to support diverse assemblages and steep increases in richness with canopy height [@oliveiraVerticalStratificationInfluences2019]. To quantify the verticality of an assemblage, we used the aforementioned trait databases, assigning verticality scores to each species based on their occurrence in fossorial (0), terrestrial (0.5), and arboreal (1) habitats [@oliveiraVerticalStratificationInfluences2019; @scheffersVerticalArborealityHorizontal2017]. For species occurring in multiple habitats, we averaged scores across habitats. We then calculated assemblage verticality as the standard effect size of the average verticality score across species in an assemblage, using 100 simulations to determine if assemblages are more vertical than expected by random chance given their richness and the evolutionary history of the biogeographic realm [@oliveiraVerticalStratificationInfluences2019] (Supplementary methods).

:::

# Vegetation structure and canopy climates impact the capacity for tall canopies to support vertical assemblages

::: {custom-style="Normal"}

The microclimate models integrate forest structure data derived  While the large computational requirements of resolving micrometeorological processes within forest canopies inhibit the use of global 3D climate data in our analysis, site-specific models spanning six regions across tropical, temperate, and boreal forests provide insight into the fine-scale processes underlying large-scale patterns.

To examine environmental factors influencing assemblage verticality, we used spatial linear mixed models to model assemblage verticality for each vertebrate class as a function of canopy height and its interaction with macroclimatic factors that may inhibit or enhance the capacity for canopies to support diverse arboreal assemblages. Macroclimate variables included those that may impact vegetation structural complexity or place physiological constraints on arboreal life strategies - minimum temperature of the coldest month, maximum temperature of the warmest month, precipitation of the driest month, and precipitation of the warmest month - as well as diurnal temperature range, which may influence the distinctness of vertical climate banding (Supplementary methods). We additionally included vegetation density in the models, which impacts habitat volume. While foliage height diversity and temperature and precipitation seasonality may also impact vertical niche opportunities and the distinctness of climate banding, these variables were strongly correlated with canopy height, minimum temperature, and dry season precipitation, respectively, and were therefore excluded from the models.

While using macroclimate data was necessary due to the lack of global vertically resolved microclimate data, macroclimate maps do not represent the microclimatic conditions experienced across vertical forest profiles that directly influence the vertical structure of animal assemblages [@defrenneWeatherStationsLack2016; @bramerAdvancesMonitoringModelling2018; @scheffersIncreasingArborealityAltitude2013]. We therefore additionally modeled vertical microclimate profiles in six 1° x 1° regions from tropical to boreal forests. To do so, we integrated information on the vertical distribution of vegetation derived from the GEDI spaceborne lidar [@dubayahGlobalEcosystemDynamics2020] with ERA5 macroclimate data at a high temporal resolution [@hersbachERA5HourlyData2023] in a process-based microclimate model that resolves within-canopy micrometeorological processes [@macleanMicroclimfFastCanopy2024; @macleanMicropointCanopyPoint2025] (Supplementary methods). By estimating the climate conditions experienced by forest-dwelling organisms, these models provide insight into the microclimatic processes that underlie macroclimate patterns and allow canopies to support vertically stratified and diverse animal assemblages.

Overall, our models of verticality indicate that taller canopies support highly vertical vertebrate assemblages, while dry and thermally variable macroclimate conditions, including extreme hot and cold temperatures limit verticality and/or the capacity for taller forests to support assemblages with higher verticality, albeit to varying degrees across vertebrate classes (Fig. \@ref(fig:fig3), S\@ref(fig:preds-univariate)-\@ref(fig:preds-interaction); Table S\@ref(tab:vert-model-results)). These patterns arise due to impacts of climate on vegetation structure and physiological constraints to arboreal strategies, as well as impacts of temporal climate variation on the distinctness of vertical climate bands.

:::

## Climate produces structural constraints to verticality

::: {custom-style="Normal"}

Climate impacts the structural complexity of vegetation, which influences the availability of arboreal niche opportunities and is typically positively associated with high animal richness [@xingEcologicalPatternsProcesses2023a; @daviesAdvancesAnimalEcology2014; @goetzLaserRemoteSensing2007]. Our models indicated that high dry season precipitation positively affects assemblage verticality across taxa (Fig. \@ref(fig:fig3)). Wet climates with low precipitation seasonality support tall and structurally complex forests with densely layered vegetation [@ehbrechtGlobalPatternsClimatic2021a; @decontoCharacterizingStructuralComplexity2024], as well as diverse epiphyte communities, which provide additional microhabitats in the canopy that contribute to high arboreal richness [@taylorVascularEpiphytesContribute2022; @thomsenSecondaryFoundationSpecies2018; @ellwoodDoublingEstimateInvertebrate2004]. Hotspots of structural complexity in wet regions that coincide with some of the tallest forests, including the Choco-Darien, the Amazon, Borneo, and New Guinea [@ehbrechtGlobalPatternsClimatic2021a; @decontoCharacterizingStructuralComplexity2024], thus provide opportunities for fine-scale niche partitioning, yielding species rich assemblages [@camargoForestVerticalComplexity2018; @gouveiaForestStructureDrives2014; @oliveiraVerticalStratificationInfluences2019].

In drier forests and higher latitudes, lower precipitation and light hinder the development of densely layered forests by restricting the functional and structural configurations of trees [@ehbrechtGlobalPatternsClimatic2021a; @hakkenbergClimateMediatesRelationship2021; @atkinsPowerLawScaling2022]. In evergreen needleleaf forests that dominate boreal regions, these environmental constraints dampen the rate of increase in structural complexity with canopy height [@atkinsPowerLawScaling2022]. Tall canopies therefore provide fewer niche opportunities in boreal regions, which may constrain assemblage verticality and limit species richness relative to canopies of equivalent height at lower latitudes. Yet, hotspots of structural complexity in temperate forests, including North Pacific Alaskan, Valdivian, and Tasmanian forests, that nearly match the complexity of tropical forests [@ehbrechtGlobalPatternsClimatic2021a] also contain assemblages with lower verticality than the tropics and fewer species than expected based upon their canopy height (Figs. \@ref(fig:fig1), \@ref(fig:fig2)). MacArthur & MacArthur [-@macarthurBirdSpeciesDiversity1961] observed this pattern over 50 years ago, noting that "tropical bird species diversity is much greater than the temperate one for habitats of comparable profiles." The interaction between macroclimate and vertical microclimate profiles provides an explanation that mirrors the climatic mechanisms allowing mountains to be better generators of biodiversity in the tropics [@rahbekHumboldtsEnigmaWhat2019].

:::

## The influence of canopy climates

::: {custom-style="Normal"}

Vertical features in the landscape, including mountains and trees bring distinct climates within close proximity of each other [@rahbekHumboldtsEnigmaWhat2019; @leahyVerticalNicheElevation2021; @vinodThermalSensitivityForest2023]. By absorbing and reflecting solar radiation and reducing vertical air mixing, dense forests produce vertical microclimate gradients that capture a substantial volume of climatic variation from the cool and wet understory to the climatically variable canopy, which experiences drier, hotter, and colder conditions [@defrenneForestMicroclimatesClimate2021]. These steep climate gradients can support high rates of species turnover from the ground to the canopy that yield vertically stratified, diverse assemblages [@bashamVerticalStratificationCollapses2020; @xingSpeciesTurnoverAnt2022; @oliveiraVerticalStratificationInfluences2019; @camargoForestVerticalComplexity2018]. However, more extreme microclimates in the canopy than the understory may place physiological constraints on arboreal lifestyles depending on background macroclimate conditions, effectively truncating the available volume of forest and limiting the capacity for tall forests to support high richness [@oliveiraVerticalStratificationInfluences2019].

Constraints of low dry season precipitation on verticality (Fig. \@ref(fig:fig3)) were accompanied by steep gradients in minimum relative humidity (Fig. \@ref(fig:fig4)). Water scarcity risks desiccation in hot or cold climates, where evapotranspiration to maintain body temperatures below thermal maxima or basking to maintain body temperatures above thermal minima can quickly lead to desiccation without adequate physiological defenses [@scheffersIncreasingArborealityAltitude2013; @tracyNotJustSmall2010; @mckechnieAvianThermoregulationHeat2017]. Dry canopy microclimates combined with low precipitation inputs therefore impose environmental filters on the functional morphologies that can inhabit canopy environments, which may limit the number of species that an increase in canopy height can support, in line with the physiological tolerance hypothesis [@curriePredictionsTestsClimatebased2004]. For example, in the seasonally dry forests of Panama, water limitation in the canopy causes the vertical stratification of frog communities to collapse toward the ground during the dry season - a shift coincident with declines in richness and homogenization of arboreal diversity [@bashamVerticalStratificationCollapses2020]. Dry climates can therefore reduce the capacity for increasingly tall canopies to support greater biodiversity (Fig. \@ref(fig:fig1)), just as cold climates limit the capacity for tall mountains to generate biodiversity by reducing the elevation of the snow line [@rahbekHumboldtsEnigmaWhat2019; @vonhumboltEssayGeographyPlants2013; @janzenWhyMountainPasses1967]. In contrast, high dry season precipitation in tropical moist forests supports highly vertical assemblages by reducing physiological constraints on arboreal life strategies in the upper canopy. Species are therefore able to finely partition the entire volume of forested habitat, leading to a dense packing of species in the canopy and allowing increases in canopy height to yield rapid increases in species richness [@oliveiraVerticalStratificationInfluences2019] (Fig. \@ref(fig:fig1)).

Maximum and minimum temperatures had less consistent effects across taxa (Fig. \@ref(fig:fig3)), but may still place physiological constraints on arboreal life strategies that influence the capacity for taller canopies to support richer species assemblages, particularly in regions exhibiting steeper vertical temperature gradients. For example, high maximum temperatures generally reduced verticality and/or reduced the positive impact of canopy height on verticality, barring mammals which exhibited the opposite response (Figs. \@ref(fig:fig3), S). High temperatures may cause ectotherm body temperatures to exceed critical thermal limits [@sundayThermalsafetyMarginsNecessity2014]. While consistent water availability in the canopies of tropical wet forests may allow amphibians to maintain body temperature below critical limits [@sundayThermalsafetyMarginsNecessity2014], the drier conditions of temperate forests combined with steeper vertical gradients in maximum temperatures may impede use of the canopy and require that animals thermoregulate by moving to terrestrial or fossorial habitats [@sundayThermalsafetyMarginsNecessity2014; @scheffersIncreasingArborealityAltitude2013; @bashamVerticalStratificationCollapses2020] (Fig. \@ref(fig:fig4)).

Moreover, low minimum temperatures reduced positive effects of canopy height on verticality for birds and reptiles despite being associated with lower verticality for mammals and amphibians due to high minimum temperatures in hot and dry regions (Fig. \@ref(fig:fig3)). The interaction between minimum temperatures and canopy height reflects physiological constraints on arboreality. Extreme cold at higher latitudes may reduce ectotherm body temperatures below critical thermal minima or impose energetic constraints on endotherms, particularly in the colder and windier conditions of the upper canopy [@sundayThermalsafetyMarginsNecessity2014; @villen-perezWoodlandMediterraneanBirds2014; @wolfThermalEffectsRadiation1996] (Fig. \@ref(fig:fig4)). To thermoregulate, animals may forage in lower canopy positions or use fossorial habitats [@sundayThermalsafetyMarginsNecessity2014; @villen-perezWoodlandMediterraneanBirds2014]. Cold temperatures therefore limit the degree to which animals can specialize to upper canopy strata and consequently prevent taller canopies from increasing species richness.

Temporal climate variability may also influence the capacity for verticality to mediate relationships between species richness and canopy height by impacting climatic distinctness. In montane regions, climatic distinctness across elevations declines with latitude due to increasing seasonal temperature variation toward the poles [@rahbekHumboldtsEnigmaWhat2019; @janzenWhyMountainPasses1967]. Low seasonal variation in the tropics produces narrow and distinct climate bands that exhibit little overlap across elevations. Unique climates across elevations restrict species' range size, leading to high rates of turnover over short distances that yield extraordinary levels of biodiversity in tropical montane regions [@janzenWhyMountainPasses1967; @sheldonFiftyYearsMountain2018; @ghalamborAreMountainPasses2006; @rahbekHumboldtsEnigmaWhat2019]. In contrast, greater temperature seasonality in temperate regions increases the similarity between climates experienced at high and low elevations, reducing the capacity for temperate mountains to support high rates of turnover and thus high levels of biodiversity [@rahbekHumboldtsEnigmaWhat2019; @ghalamborAreMountainPasses2006].

In forests, diurnal temperature range may influence the distinctness of vertical climate bands from the ground to the canopy, and thus the vertical turnover of species. Our models of verticality indicate that greater diurnal temperature variation reduces verticality of mammals, though effects for other classes were insignificant (Fig. \@ref(fig:fig3)). Furthermore, microclimate models indicated that lower daily temperature variation and greater decoupling between canopy and understory temperatures generally reduces thermal overlap across vertical strata in tropical forests relative to temperate and boreal forests (Fig. \@ref(fig:fig4)b, S\@ref(fig:microtempdiu)). Without unique climatic niches across the forest profile, high latitude forests are unlikely to support rapid vertical turnover, limiting the number of species that may be added to the assemblage for a given increase in canopy height, even in wetter temperate regions with climate conditions that physiologically support arboreal strategies. In contrast, tropical canopies maintained more distinct vertical climate bands throughout the day, which may support rapid vertical turnover, leading to greater increases in richness with canopy height and enhancing the capacity for tropical canopies generate biodiversity.

[TRY LABELLING MICROCLIM OVERLAP FIG AS BOREAL, TEMPERATE, TROPICAL COLUMNS AND EASTERN AND WETERN
HEMISPHERE ROWS TO SAVE SPACE ALSO REORDE PROFILE PANELS INTO THE ORDER THEY ARE DISCUSSED]

:::

# Canopy height in a changing world

::: {custom-style="Normal"}

Centuries of research have elucidated the mechanisms that allow mountains to generate high levels of biodiversity, particularly in the tropics [@vonhumboltEssayGeographyPlants2013; @janzenWhyMountainPasses1967; @rahbekHumboldtsEnigmaWhat2019]. Yet only recently has the three-dimensional nature of terrestrial landscapes been widely recognized as a critical dimension to our understanding of biogeographic patterns [@fredstonReimaginingSpeciesMove2025; @xingEcologicalPatternsProcesses2023a]. We bring the vertical dimension of habitats to the forefront of global biogeography by showing that the climatic characteristics of wooded ecosystems from macro- to micro-scales mediate the capacity for tall forests to support high levels of biodiversity via impacts on the vertical structure of vertebrate assemblages. Tropical rainforests, with wet and thermally stable climates that increase the structural complexity of vegetation, reduce physiological stress in the canopy, and produce more distinct vertical climate bands, can support fine-scale partitioning of resources and climatic niches throughout the entire forest volume. Consequently, taller trees in tropical forests support higher biodiversity and greater increases in species richness than temperate and boreal ecosystems of equivalent heights where climatic conditions reduce the structural complexity of vegetation, pose physiological constraints on arboreal lifestyles, and reduce vertical climate zonation from the ground to the canopy.

While tall forests, particularly in the tropics, are crucial for supporting species rich assemblages, regions containing tall forests are globally rare, under-protected, and threatened by anthropogenic disturbance and climate change [@huangPrioritizingGlobalTall2023; @bourgoinHumanDegradationTropical2024; @bennettLargerTreesSuffer2015; @rowlandDeathDroughtTropical2015]. Disturbances from fire and logging produce persistent reductions in canopy height in the tropics [@bourgoinHumanDegradationTropical2024], with potential negative effects on the capacity for these forests to support high levels of biodiversity. Furthermore, climate change is increasing the frequency of extreme heat and drought [@seneviratneWeatherClimateExtreme2021]. In addition to directly reducing forest height [@bennettLargerTreesSuffer2015; @rowlandDeathDroughtTropical2015], these trends could truncate the volume of physiologically suitable habitat for arboreal species. Climate change may therefore flatten the vertical structure of species assemblages toward the ground [@oliveiraVerticalStratificationInfluences2019; @scheffersIncreasingArborealityAltitude2013; @bashamVerticalStratificationCollapses2020; @soiferMultidimensionalMicroclimateVelocities2025], just as warming temperatures are causing upslope shifts in species' distributions [@lenoirSpeciesBetterTrack2020; @chanClimateVelocitiesSpecies2024]. Improving our understanding of the vertical structure of species assemblages as canopies heat up and dry out will be critical for evaluating the continued capacity of tropical forests to generate extraordinarily high levels of biodiversity.

:::

# Data availability statement

canopy height:

# Acknowledgements
LGS was supported by the NSF GRFP...
We thank Patrick Burns for access to the processed GEDI L2b points and Alex Baecher for assistance with the spatial statistics.

\newpage

# Figures

```{r fig1, fig.id="fig1", fig.cap="Spatial distribution of a) canopy height, b) vertebrate richness (for all vertebrates), and c) deviance residuals for a negative binomial model of total vertebrate richness by canopy height. d) Relationships between species richness of all vertebrates and canopy height across biomes. Dots indicate observed values and lines indicate negative binomial model fits of richness as a function of the interaction between canopy height and biome.", fig.height=100, fig.width=140}

# Fig. 1

include_graphics("figures/ms_figures/fig1-2_canopy_richness.png", error = F)

```

```{r fig2, fig.id="fig2", fig.cap="Global variation in the vertical structure of vertebrate assemblages and its association with species richness for birds, mammals, reptiles, and amphibians. a) The average ± SD proportion of species exhibiting arboreal (red), terrestrial (yellow), and fossorial (blue) strategies across latitudes. Proportions can exceed one because species can exhibit multiple vertical strategies. b) The verticality of vertebrate communities in wooded biomes across the globe. High verticality (red) indicates that the assemblage has a higher vertical niche position than expected by random chance. c) Correlations between verticality and species richness globally. Lines indicate linear best fit line with 95% CIs.", fig.width = 150, fig.height=150}

# Fig. 2

include_graphics("figures/ms_figures/fig2_vertmaps.png", error = F)

```

```{r fig3, fig.id="fig3", fig.cap="Coefficients of standardized predictors for spatial LMMs of verticality as a function of environmental factors. Models were conducted for each vertebrate class. Error bars represent 95% confidence intervals. Filled circles with solid error bars have confidence intervals that do not overlap zero. Open circles with dashed error bars have confidence intervals that overlap zero. Tmin~cold~ = minimum temperature of the coldest month; Tmax~warm~ = maximum temperature of the warmest month; Temp~diu~ = diurnal temperature range; Precip~dry~ = precipitation of the driest quarter; Precip~warm~ = precipitation of the warmest quarter.", fig.width=140, fig.height=120}

# Fig. 3

include_graphics("figures/ms_figures/fig3_sdmTMB_coefs.png", error = F)

```

THINGS TO FIX IN FIG4 - diurnal temp range is for warmest month in strata plot and coldest month in
profile plot - make consistent - replace map with concept figure similar to Rahbek - should we add
strata plots for other variables? Or add to supplement - probable supplement - text sizing - point
size in map if we keep the map

```{r fig4, fig.id="fig4", fig.cap="a) Locations of modeled vertical microclimate profiles. b) Average hourly microclimate across relative heights in the canopy. Relative height is calculated as absolute height above ground divided height of the canopy. Shading represents the range of temperatures experienced in the uppermost vertical band (0.9-1 relative height) and the lowermost vertical band (0-0.1 relative height). c) Vertical profiles of microclimate variables and plant area volume density (PAVD) (median +/- SD) in tropical, temperate, and boreal forest sites. PAVD was obtained from GEDI shots.", fig.width=180, fig.height=110}

# Fig. 4

include_graphics("figures/ms_figures/fig4_profiles.png", error = F)

```

# Supplementary material

# Materials and Methods

## Assemblage richness and verticality

```{r}
amph_spdat = read.csv("data/derivative_data/species_data/amph_spdat_moura_forestsOnly.csv")
rept_spdat = read.csv("data/derivative_data/species_data/reptiles_spdat_moura_forestsOnly.csv")
mammals_spdat = read.csv("data/derivative_data/species_data/mammals_spdat_moura_forestsOnly.csv")
bird_spdat = read.csv("data/derivative_data/species_data/birds_spdat_elton_forestsOnly.csv")

namph = nrow(amph_spdat)
nrept = nrow(rept_spdat)
nmammals = nrow(mammals_spdat)
nbirds = nrow(bird_spdat)

```

We mapped global richness and verticality of terrestrial vertebrate assemblages in wooded habitats [@dinersteinEcoregionBasedApproachProtecting2017] within 50-km grid cells using a cylindrical equal area projection. Range maps were obtained from the IUCN Red List for mammals and amphibians [@iucnIUCNRedList2023], BirdLife International for bird breeding and resident ranges [@birdlifeinternationalandhandbookofthebirdsoftheworldBirdSpeciesDistribution2022], and the Global Assessment of Reptile Distributions (GARD) for reptiles [@rollGlobalDistributionTetrapods2017; @caetanoAutomatedAssessmentReveals2022]. Overall, we obtained range maps for `r nbirds` birds, `r nmammals` mammals, `r nrept` reptiles, and `r namph` amphibians. Wooded habitats included  all forested biomes, as well as ecoregions identified as forests, pine,  woodlands, woods, taiga, espinal, várzea, yungas, savanna, thicket, matorral,  chaparral, mezquital, pineyards, pantanal, the cuban wetlands, cerrado, bushveld, chaco, lowveld, and mallee as defined by Dinnerstein et al. [-@dinersteinEcoregionBasedApproachProtecting2017]. We did not include mangroves given their lack of terrestrial and fossorial habitat.

To calculate the verticality of each assemblage (i.e., grid cell), we assigned each species a verticality score from zero to one and then took the average verticality of all species whose ranges overlapped with the grid cell. For mammals, reptiles, and amphibians, verticality scores were based on their use of fossorial (0), terrestrial (0.5), and arboreal (1) habitats as defined in the TetrapodTraits database [@mouraPhylogenyinformedCharacterisationGlobal2024a]. The TetrapodTraits dataset uses phylogenetic imputation to identify traits if trait data is not available in the literature. For species where imputation was necessary, we considered a species to use a given vertical stratum if over 70% of the 10,000 imputations predicted use of the stratum. For birds, we calculated verticality scores based on their use of different foraging strata as identified in the EltonTraits database [@wilmanEltonTraitsSpecieslevelForaging2014], which provides a finer scale vertical classification than the dataset from Moura et al. [-@mouraPhylogenyinformedCharacterisationGlobal2024a]. For birds, verticality scores were as follows: foraging below the water surface, around the water surface, and on the ground = 0.5, foraging in the understory = 0.667; foraging in the mid-high canopy = 0.8337; foraging in or above the upper canopy = 1. For all taxa, if species used multiple vertical habitats, vertical scores were averaged. For example, if a species occurred in terrestrial (0.5) and arboreal (1) habitats, it would receive a verticality score of 0.75. Species that occupied only aerial vertical niches or that did not occur in terrestrial habitats (e.g., marine mammals) were excluded from the analysis.

To examine spatial patterns of assemblage verticality independently of species richness and accounting for biogeographic history, we calculated the standard effect size (SES) of mean verticality (hereafter ‘verticality’): $SES=(x_i-\bar{x})/sd(x)$ where x is a vector of expected mean verticality scores given random selection of species from a species pool. For each vertebrate class, we defined species pools as all species inhabiting wooded habitats within a given biogeographic realm [@dinersteinEcoregionBasedApproachProtecting2017] and generated a null distribution of expected verticality for each grid cell based on 100 simulations. Realms included Palearctic, Nearctic, Indomalayan, Neotropic, Oceania, Afrotropic, and Australasia. For each simulation, we randomized verticality within the grid cell by randomly drawing the same number of verticality scores from the respective species pool as the observed richness in the cell. Highly positive SES verticality scores represent assemblages having higher verticality than expected given the species richness, while low SES verticality scores represent assemblages having much lower verticality than expected given the species richness.

## Environmental data

We obtained macroclimate data from CHELSA v2.1 (30 arc-second resolution), representing average conditions for the time period 1981-2010 [@kargerDataClimatologiesHigh2021; @kargerClimatologiesHighResolution2017]. Macroclimate variables represented climatic conditions that could pose physiological constraints to arboreal life styles or impact the distinctness of vertical climate banding, and included mean daily maximum air temperature of the warmest month (Tmax~warm~; °C), mean daily minimum air temperature of the coldest month (Tmin~cold~; °C), precipitation amount of the driest month (Precip~dry~; kg m^-2^ month^-1^), mean monthly precipitation of the warmest quarter (Precip~warm~; kg m^-2^ month^-1^), temperature seasonality (°C), mean diurnal temperature range (Temp~diu~; °C), and precipitation seasonality (Precip~sea~; kg m^-2^).

We obtained vegetation data, including canopy height [@langHighresolutionCanopyHeight2023], foliage height diversity [@burnsMultiresolutionGriddedMaps2024] vegetation density [@crowtherMappingTreeDensity2015]. The canopy height layer represents canopy height at a 10 m resolution for the year 2020 and was developed using a fusion of height data from the Global Ecosystem Dynamics Investigation (GEDI) and optical satellite images from Sentinel-2. To obtain canopy height in in-tact wooded habitat, excluding human modified land covers, such as plantations and shifting agriculture, we applied a mask using Hansen et al. [-@hansenGlobalLandUse2022] global land cover and land use in 2019. To do so, we aggregated the data to match the resolution of the land cover layer (~ 30 m), and then masked all grid cells with woody vegetation \< 3 m, surface water and permanent ice, built-up land, cropland, recent forest loss, and land use associated with forestry and shifting cultivation (e.g., oil palm). Foliage height diversity (FHD) was obtained from gridded GEDI data at a 1 km resolution and represents the vertical heterogeneity of vegetation. FHD was masked using the same method as canopy height. The vegetation density map represents stems per hectare and was generated using regression models linking stem density in forest plots to climate, topography, vegetation characteristics, and the proportion of developed land cover. Because the vegetation density estimates were generated using plots in forests rather than spatially continuous observations including human modified land use types, we did not mask vegetation density with the land cover map.

All environmental data were resampled to the spatial resolution and projection of the species data.

## Vertebrate richness models

Vertebrate richness is expected to increase with increasing canopy height. We first modeled richness as a function of canopy height using a generalized linear model with a Poisson distribution and a log-link. Models were produced for global richness across all vertebrate classes and global richness for each vertebrate class. For each vertebrate class, we then modeled richness as a function of the interactive effect between canopy height and biome using a means parameterization. Initial model evaluation indicated over-dispersion in model residuals. We therefore fit negative binomial models with a log-link function. Since richness often exhibits a log-log relationship with covariates, we additionally fit models of richness as a function of log(canopy height) or the interactive effect of log(canopy height) and biome, and compared this model to the original using AIC and conditional percent deviance explained (CPDE). Models fit with canopy height rather than log(canopy height) consistently performed better, and we report these results.

We classified biomes into larger categories based on the [@dinersteinEcoregionBasedApproachProtecting2017] 2017 biome categories. For biomes that occurred across latitudes (e.g., Montane Grasslands and Shrublands), we classified them as tropical if they occurred within the latitudes covered by biomes classified as tropical and temperate if they occurred within the latitudes classified as temperate. Additionally, Mediterranean forests, woodlands, and scrub were partitioned into temperate forests if the ecoregion contained "forest" and temperate woodland, savanna, and shrubland if the ecoregion name did not contain the word "forest".

We additionally examined relationships between richness and verticality using Pearson correlations.

## Verticality models

We examined relationships between the verticality of each taxon and environmental covariates described above using spatial linear mixed models (LMMs). Grid cells with fewer than five species were excluded, because they may provide unreliable estimates of the SES. To linearize the relationships between predictor and response variables, we log10-transformed precipitation of the driest month. We then scaled all predictor variables to mean zero and variance one and examined their multicollinearity using Pearson correlations and variance inflation factors. We excluded precipitation seasonality, temperature seasonality, and foliage height diversity due to strong correlations (|r| > 0.8) with dry season precipitation, minimum temperature of the coldest month, and canopy height, respectively. The remaining variables all had VIF < 5. Additionally, we included a quadratic effect for maximum temperature of the warmest month, because extremely high temperatures are expected to inhibit arboreal life strategies as well as interaction effects between the climatic covariates and canopy height. Interactions were included because we expect the impact of canopy height on verticality to decline when harsh macroclimate conditions, including extreme cold, hot, or dry conditions, reduce the suitability of canopy microhabitats for arboreal life strategies, or when high climate variability reduces distinctness of vertical climate bands. The fixed effects therefore included Tmax~warm~, Tmax~warm~^2^, Tmin~cold~, Temp~diu~, Precip~warm~, log10(Precip~dry~), vegetation density, and canopy height, as well as interactions between canopy height and the climatic covariates.

Spatial LMMs were fit using the R package sdmTMB [@andersonSdmTMBPackageFast2024], which accounts for spatially autocorrelated data using stochastic partial differential equations (SPDE) to approximate spatial Gaussian random fields as Gaussian Markov Random Fields (GMRF) with a Matérn covariance function [@lindgrenExplicitLinkGaussian2011]. The full model can be described as:

$$
\begin{align}
y_s ∼ N(𝜇_{s}, \sigma^2), \\
𝜇_s = 𝜲_s𝛽 + ⍵_s, \\
⍵_s ∼ MVNormal(0,𝚺_⍵),
\end{align}
$$

where the response variable at location $s$ is normally distributed with mean $\mu_s$ and variance $\sigma^2$; $𝜇_s$ is the expected verticality. $Χ_s$ is a design matrix at location $s$; 𝛽 is a vector of fixed-effect coefficients; $⍵_s$ represents a gaussian spatial random field at location $s$; and $\Sigma_{\omega}$ represents the covariance matrix of the spatial random field. The SPDE method requires a finite element analysis mesh, which consists of a set of triangles covering the spatial domain, and allows for the approximation of the spatial random effect at mesh vertices [@thorsonSpatioTemporalModelsEcologists2024]. The mesh used to approximate the spatial random field impacts model results as well as computation time [@andersonSdmTMBPackageFast2024]. To produce an optimal mesh that would avoid over-smoothing the spatial random field and avoid over-fitting the model and unnecessarily increasing computational demands, we used an iterative procedure to update minimum and maximum edge lengths of the triangles in the mesh based on the spatial range of the data (i.e., the distance at which points exhibit a correlation of approximately 0.1) [@thorsonSpatioTemporalModelsEcologists2024]. The tuning procedure for the mesh was conducted for each taxon using *fmesher* and consisted of the following steps. First, we identified the spatial range of the verticality using a correlogram on a subset of 1000 points. Next, we constructed an initial mesh with a maximum edge length of the inner domain as 1/5 the spatial range, an inner extension equal to the length of the spatial range, an outer extension equal to twice the length of the spatial range, and the maximum edge length in the outer extension equal to the spatial range. To avoid small triangles, we also specified the minimum edge lenght as 1/5 the maximum edge length of the inner domain [@krainskiAdvancedSpatialModeling2019]. We then fit a model using the resulting mesh. If the estimated range in the model was less than five times the maximum edge length of the spatial domain used to construct the mesh, we constructed a new mesh, updating the edge lengths and buffers based on the spatial range estimated by the model. We repeated this procedure until the spatial range estimated by the model was greater than 5 times the maximum edge length of the mesh.

We evaluated models using the sanity check function in sdmTMB and by examining quantile residuals using the DHARMa R package [@hartigDHARMaResidualDiagnostics2022]. To determine the degree to which environmental factors improve model performance, we compare the models to spatial random effects models without environmental predictors and report AIC and conditional percent deviance explained (CPDE).

## Microclimate models

To examine vertical microclimate gradients, we mechanistically modeled microclimates from the ground to the canopy using the ‘micropoint’ R package [@macleanMicropointCanopyPoint2025; @macleanMicroclimfFastModelling2025]. Point models of vertical climate profiles were produced for boreal, temperate, and tropical forests from the Americas and Eurasia within six 1° x 1° tiles located in regions with relatively flat topography that contained relatively large areas of tree cover based on Google satellite imagery and that were within the latitudinal bounds of the GEDI data, which provided information on vegetation structure (see below). Boreal forests included regions in Canada (50°,-90°) and Russia (50°,137°); temperate forests included regions in the midwestern United States (37°,-92°) and Europe (48°,10°); tropical forests included regions in northern Peru (-6°,-76°) and Malaysia (4°,117°), where coordinates indicate the lower left corners of tiles.

Points within these tiles were selected as locations of GEDI footprint level L2b data, which was preprocessed by Burns et al. [-@burnsGriddedGEDIVegetation2024; @burnsMultiresolutionGriddedMaps2024] to filter out low quality points. From the high quality points, we excluded points that 1) fell outside closed forest, as identified by Copernicus Global Dynamic Land Cover 2019 v3 [@buchhornCopernicusGlobalLand2020], 2) were located in areas that experienced deforestation between 2001 and 2024, as identified by Hansen global forest watch [@hansenHighResolutionGlobalMaps2013], 3) were obtained from coverage rather than power beams, 4) represented vegetation less than 10 m tall, and 4) occurred above 500 m in elevation. We restricted analysis to low elevation forests to control for any elevational impacts on vertical climate gradients. From the remaining points, we identified those that were recorded during the months with the average coldest minimum temperature and average warmest maximum temperature for the year in which they were obtained. These months were identified using the ECMWF ERA5 reanalysis product [@hersbachERA5HourlyData2023]. For minimum and maximum temperature subsets, we then selected up to 100 GEDI points within each 1° x 1° tile from the top 75th to 98th quantile of canopy height. However, there were only 26 points in temperate USA and 14 points in temperate Europe during minimum temperature months because GEDI only produces reliable vertical profiles during leaf-on conditions (cite). The points retained therefore represented mixed rather than deciduous forest types. Furthermore, only 65 points met our selection criteria in temperate European forests during the warmest month. For each point, the microclimate model was run for the month and year in which the point was obtained at 0.15 cm above ground, 2 m above ground, and then from 5 m to the top of the canopy at 5 m intervals (equivalent to the vertical resolution of GEDI data) to produce hourly estimates of air temperature and relative humidity.

### Climate data

The ECMWF ERA5 reanalysis product [@hersbachERA5HourlyData2023] was used for meteorological inputs to the model, which provided near surface, open-air climate conditions at a 0.25$^\circ$ spatial resolution and hourly temporal resolution. ERA5 climate data was processed using the *mcera5* R package, which converts ERA5 variables into those required by the model and applies a coastal diurnal temperature correction to terrestrial locations located within predominantly marine grid cells [@klingesMcera5DrivingMicroclimate2022]. Meteorological conditions at the GEDI point of interest were then extracted from the climate grid by taking the weighted average of climate conditions across the four nearest neighboring cells, where the weight was defined as the inverse distance to the point. Finally, we applied an elevational correction that adjusts temperature and pressure at the location of interest by accounting for the difference in elevation between the center of the ERA5 grid cell and the point of interest using a humidity-dependent lapse rate.

### Vegetation data

Vegetation structure influences the amount of radiation and vertical air mixing. We obtained information on vegetation from the GEDI footprint level L2b data.
For each point, GEDI data provided information on canopy height (RH100; the height at which 100% of the waveform was returned), plant area index (PAI), and the vertical distribution of the PAI. The GEDI data describes the vertical distribution of PAI as the cumulative PAI from the top of the canopy to the stratum representing the vertical space from 0-5 m above the ground. For example, the PAI 0-5 m includes nearly all of the canopy space and is therefore nearly equal to the total PAI in most cases, while PAI 10-15 m describes the PAI from the top of the canopy to 10-15 m above the ground. However, the microclimate models require that the vertical PAI distribution represent the PAI within individual height bins and that the sum of the vertical PAI profile equal the total PAI. We therefore subtracted total PAI – PAI 0-5 m to represent the PAI in the 0-5 m bin, PAI 0-5 m – PAI 5-10 m to represent the PAI in the 5-10 m bin, and so forth.

*Micropoint* requires a number of additional parameters describing vegetation structure that are not available from remote sensing products, such leaf angle distribution and stomatal conductance, among others. We use *microclimf's* 'vegpfromhab()' function to estimate these parameters based on land cover type. Land cover was determined based on MODIS land cover type (MCD12Q1) version 6.1 for the year 2022.

Vegetation parameters were assumed to remain constant for the month and year in which the GEDI shot was obtained.

[WE SHOULD PRODUCE MODELS AT DRIEST MONTH OF THE YEAR]

### Topography and soil composition

To characterize topography, we extracted elevation from the lowest mode return of the GEDI shot. Slope and aspect were calculated using *terrain()* based on SRTM elevation data at a 1 km resolution, which was obtained using the *geodata* R package. Sand, silt, and clay content in the top soil layer (0-5 cm below the ground) were obtained from the SoilGrids 2.0 at a 250 m resolution using the soilDB R package, and included sand, silt, and clay content at the top soil layer (0-5 cm below the ground). Soil type was determined based on the relative percentage of sand, silt, and clay based on USDA soil classifications using the *soiltexture* R package. Soil properties were then defined for the given soil class as default values provided by *micropoint*.

### Microclimate analysis

To examine global patterns of vertical microclimate profiles, we calculated the average minimum temperature of the coldest month, maximum temperature of the warmest month, minimum relative humidity of the warmest month, and diurnal temperature range of the coldest month. For each forest type (boreal, temperate, and tropical), we graphed the median climate value within vertical intervals, each representing 10% of the total canopy height. Therefore, values at the top of the canopy always occur in the 0.9-1 relative height interval regardless of the absolute height of the canopy. We additionally examined diurnal temperature fluctuations across forest types by taking the mean value across points for a given relative height and hour of the day.

<!---BLOCK_LANDSCAPE_START--->

# Supplementary tables

```{r richness-mod-results, tab.id="richness-mod-results", tab.lp="stab", tab.cap="Coefficient estimates (with 95% confidence intervals) and conditional percent deviance explained (CPDE) for global models of richness as a function of canopy height for all taxa and each vertebrate class and for biome models of richness as a fucntion of canopy height x biome. Biome models used a means parameterization such that the effect of biome represents the mean estimated richness at a canopy height of zero and the interaction effect represents the estimated impact of canopy height on richness for each biome."}

# tables from: "scripts/03_analyses/01_canopyheight_richness_relationships.R"

tidy.global = fit.global.tidy %>% 
  mutate(term = ifelse(term == "canopy_height2", "Canopy height", term),
         taxa = "All vertebrates") %>% 
  mutate(across(.cols = estimate:conf.high, .fns = function(x) {round(x,2)})) %>% 
  mutate(estimate = ifelse(term != "CPDE (%)",
                           paste0(estimate, " (", conf.low, ", ", conf.high, ")"),
                           paste0(estimate))) %>% 
  dplyr::select(taxa, term, estimate) %>% 
  pivot_wider(names_from = "taxa", values_from = "estimate")

tidy.global.taxa = global.taxa.tidy %>% 
  dplyr::select(taxa, term, estimate, conf.low, conf.high) %>% 
  mutate(term = ifelse(term == "canopy_height2", "Canopy height", term)) %>%
  mutate(across(.cols = estimate:conf.high, .fns = function(x){round(x,3)})) %>% 
  arrange(taxa) %>% 
  mutate(estimate = ifelse(term != "CPDE (%)",
                           paste0(estimate, " (", conf.low, ", ", conf.high, ")"),
                           paste0(estimate)),
         taxa = factor(taxa, levels = c("Birds", "Mammals", "Reptiles", "Amphibians"))) %>% 
  dplyr::select(taxa, term, estimate) %>% 
  pivot_wider(names_from = "taxa", values_from = "estimate")

tidy.global = left_join(tidy.global, tidy.global.taxa, by = "term") %>% 
  mutate(model = "Global model") %>% 
  relocate(model)

tidy.biome = biome.mod.tidy %>% 
  dplyr::select(taxa, term, estimate, conf.low, conf.high) %>% 
  mutate(term = gsub("e2", "e = ", term),
         term = gsub(":", " x ", term),
         term = gsub("2", "", term)) %>% 
  mutate(across(.cols = estimate:conf.high, .fns = function(x){round(x,3)})) %>% 
  mutate(estimate = ifelse(term != "CPDE (%)",
                           paste0(estimate, " (", conf.low, ", ", conf.high, ")"),
                           paste0(estimate)),
         taxa = factor(taxa, levels = c("Birds", "Mammals", "Reptiles", "Amphibians"))) %>% 
  dplyr::select(taxa, term, estimate) %>% 
  pivot_wider(names_from = "taxa", values_from = "estimate") %>% 
  mutate(model = "Biome model") %>% 
  relocate(model)

tidymods = bind_rows(tidy.global, tidy.biome)

tidymods %>% 
  flextable() %>% 
  hline(i = 3, part = "body", border = fp_border(color = "black", width = 3)) %>% 
  merge_v(j = c("model")) %>%
  set_table_properties(width = 1, layout = "autofit")

```

```{r vert-model-eval, tab.id="vert-model-eval", tab.lp="stab", tab.cap="Degrees of freedom, AIC, and conditional percent deviance explained (CPDE) for models of verticality by environmental predictors and spatial random effects and models of verticality by spatial random effects only. In the models, 'Fixed' refers to fixed environmental predictors and 'Spatial' refers to spatial random effects."}

tab = read.csv("scripts/03_analyses/03_glmms/result_tables/compareMods.csv")

tab %>% 
  mutate(across(.cols = AIC:CPDE, function(x) {round(x,2)})) %>% 
  rename(Model = name, "CPDE (%)" = "CPDE", "\u0394AIC" = "deltaAIC") %>% 
  flextable() %>% 
  merge_v(j = c("taxon")) %>% 
  bg(i = c(1:3,7:9), j = 1:6, bg="grey80") %>% 
  fix_border_issues() %>% 
  autofit()

```

```{r vert-model-results, tab.id="vert-model-results", tab.lp="stab", tab.cap="Coefficient estimates (with 95% confidence intervals) for models of verticality as a function of standardized predictor variables."}

# tab produced in "scripts/03_analyses/03_glmms/04_resultsTab.R"
tab = read.csv("scripts/03_analyses/03_glmms/result_tables/modelCoefs.csv")

tab %>%   flextable() %>% 
  compose(i = ~term == "tmax_warm", j = "term", value = as_paragraph("Tmax", as_sub("warm"))) %>% 
  compose(i = ~term == "I(tmax_warm^2)", j = "term", value = as_paragraph("Tmax", as_sub("warm"), as_sup("2"))) %>% 
  compose(i = ~term == "tmin_cold", j = "term", value = as_paragraph("Tmin", as_sub("cold"))) %>% 
  compose(i = ~term == "temp_diu", j = "term", value = as_paragraph("Temp", as_sub("diu"))) %>% 
  compose(i = ~term == "precip_warm", j = "term", value = as_paragraph("Precip", as_sub("warm"))) %>% 
  compose(i = ~term == "log_precip_dry", j = "term", value = as_paragraph("log10(precip", as_sub("dry"), ")")) %>% 
  compose(i = ~term == "canopy_height2", j = "term", value = as_paragraph("Canopy height")) %>% 
  compose(i = ~term == "veg_den", j = "term", value = as_paragraph("Vegetation density")) %>% 
  compose(i = ~term == "precip_warm:canopy_height2", j = "term", value = as_paragraph("Precip", as_sub("warm"), ":Canopy height")) %>% 
  compose(i = ~term == "tmin_cold:canopy_height2", j = "term", value = as_paragraph("Tmin", as_sub("cold"), ":Canopy height")) %>% 
  compose(i = ~term == "tmax_warm:canopy_height2", j = "term", value = as_paragraph("Tmax", as_sub("warm"), ":Canopy height")) %>% 
  compose(i = ~term == "I(tmax_warm^2):canopy_height2", j = "term", value = as_paragraph("Tmax", as_sub("warm"), as_sup("2"), ":Canopy height")) %>% 
  compose(i = ~term == "log_precip_dry:canopy_height2", j = "term", value = as_paragraph("log10(precip", as_sub("dry"), "):Canopy height")) %>% 
  compose(i = ~term == "temp_diu:canopy_height2", j = "term", value = as_paragraph("Temp", as_sub("diu"), ":Canopy height")) %>% 
  autofit()

```

<!---BLOCK_LANDSCAPE_STOP--->

\newpage

# Supplementary figures

```{r splot-global-chrich, fig.id="global-chrich",  fig.lp="suppfig", fig.autonum.start_at=1, fig.cap="Global relationship between total vertebrate richness and canopy height. Line indicates predictions from a negative binomial model of richness by canopy height ± 95% CI. Points indicate observed richness by canopy height.", , fig.width=90, fig.height=90}

include_graphics("figures/ms_figures/supp_figs/total_richness~canopy.png", error = F)
```

```{r splot-resids-taxa, fig.id="resids-taxa", fig.lp="suppfig", fig.cap="Deviance residuals from negative binomial models of richness by canopy height for each vertebrate class: a) birds, b) mammals, c) reptiles, d) amphibians.", fig.width=180, fig.height=100}

include_graphics("figures/ms_figures/supp_figs/deviance_resids_taxa.png", error = F)
```

```{r splot-meanvert-biome, fig.id = "meanvert-biome", fig.lp="suppfig", fig.cap="Distribution of verticality within a) biomes as classified by Dinerstein et al. (2017) and b) ecoregions aggregated to coarser classifications. Red diamonds indicate mean assemblage verticality. Ecoregions that do not contain woody vegetation were excluded.", fig.width = 180, fig.height = 130}

include_graphics("figures/ms_figures/supp_figs/biome_mean_verticality.png", error = F)
```

```{r splot-meanprop-biome, fig.id="meanprop-biome", fig.lp="suppfig", fig.cap="Distribution of verticality within a) biomes as classified by Dinerstein et al. (2017) and b) ecoregions aggregated to coarser classifications. Red diamonds indicate mean assemblage verticality. Ecoregions that do not contain woody vegetation were excluded.", fig.width = 160, fig.height = 160}

include_graphics("figures/ms_figures/supp_figs/biome_mean_proportion_strategies.png", error = F)
```

```{r splot-preds-univariate, fig.id="preds-univariate", fig.lp="suppfig", fig.cap="Conditional plots of model predictions for assemblage verticality of each vertebrate class in response to different environmental predictors. All variables except for that displayed on the x-axis were held constant at their mean. Error bars represent 95% confidence intervals.", fig.width=180, fig.height=120}

include_graphics("figures/ms_figures/supp_figs/preds_univariate.png", error = F)

```

```{r splot-preds-interaction, fig.id="preds-interaction", fig.lp="suppfig", fig.cap="Conditional plots of model predictions for assemblage verticality of each vertebrate class across canopy heights and at different values of climatic interaction terms. Interaction terms were allowed to vary across their mean, -2 standard deviations, and +2 standard deviations, while all other predictions were held constant at their mean. Error bars represent 95% confidence intervals.", fig.width=180, fig.height=120}

include_graphics("figures/ms_figures/supp_figs/preds_interaction.png", error = F)

```

```{r splot-microtempdiu, fig.id="microtempdiu", fig.lp="suppfig", fig.cap="Average hourly microclimate across relative heights in the canopy. Relative height is calculated as absolute height above ground divided height of the canopy.", fig.width = 120, fig.height = 70}

include_graphics("figures/ms_figures/supp_figs/microclim_temp_diu_mintempMonth.png", error = F)

```

\newpage

# References